### 2012-12-02 project/report/paper.tex
report first pass, experimental results done, preliminary outline, committing pngs, hopefully we have room

% For tracking purposes - this is V3.1SP - APRIL 2009
\documentclass{acm_proc_article-sp}
\usepackage{graphicx}
\usepackage{subfig} % for horizontal figures
\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\begin{document}

\title{OL-LMS}

\numberofauthors{2}
\author{
\alignauthor
Forrest N. Iandola\\
       \affaddr{University of California, Berkeley, USA}\\
       \email{forresti@eecs.berkeley.edu}
\alignauthor
Huy Vo\\
       \affaddr{University of California, Berkeley, USA}\\
       \email{huytbvo@eecs.berkeley.edu}
}

\maketitle
\begin{abstract}
Best abstract ever
\end{abstract}

\keywords{embedded systems and planes and stuff}

\section{Introduction}
Overview, stuff about platform based design...
\subsection{OL-LMS Goals}
Talk about goals of project
\section{HL-LMS}
Reference Mehdi's et al work on HL-LMS. Brief description to show that we understand it.
\section{LL-LMS}
Talk about our LL-LMS system, its inputs/outputs, functionality, etc
\section{OL-LMS}
Talk about how the entire system works at a whole, how we set it up, etc.
\section{Experimental Results}
\begin{figure}[hb]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/preqnofail.png}
  \caption{\textbf{Power Request}
  Historical data of power request on Bus 1 and Bus 2 for time 0
  to 100. Notice the spikes in power usage on Bus 2 that pulls the
  request over 100000 Watts, well over the power rating of the generators.
  Load shedding is a must during those intervals.}
  \label{fig:preqnofail}
\end{figure}

\subsection{No Failures}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/lshlnofail.png}
  \caption{\textbf{HL-LMS Load Shedding} Notice the long
  period of load shedding around the power usage spikes.}
  \label{fig:lshlnofail}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/lsllnofail.png}
  \caption{\textbf{LL-LMS Load Shedding} Notice that the 
  LL-LMS sheds much more load than the HL-LMS.}
  \label{fig:lsllnofail}
\end{figure}

For our first experiment, we simulated the case when there are no power
generators failures in the system using the power usage profile shown in
Figure~\ref{fig:preqnofail}. Notice that Bus 1 uses more than 100000 Watts
between time 30 and 60 and again around time 90. Bus 2, on the other hand,
has a massive power usage between time 50 and 80. We wanted to see how the
HL-LMS and the LL-LMS responds to this scenario in isolation. Our results
are in Figure~\ref{fig:lshlnofail} and Figure~\ref{fig:lsllnofail}.
Comparing the results, we can see that the LL-LMS will shed a much larger
number of loads than the the HL-LMS. In particular, the LL-LMS sheds
exclusively from load 6-10 because it simply examines the priority table.
The HL-LMS; however, is much smarter. It is able to minimize the number of
loads shed by shedding loads much higher up in the priority table.

\subsection{Generator Failure}
For this experiment, we wanted to see how a combined HL-LMS and LL-LLMS
system would respond to spontaneous generator failures. To this end, we
simulated the case when generator 2 would fail at time 45. Obviously, if the HL-LMS were
to run in isolation, it would leave the system unpowered from time 45 to 50. 
Figure~\ref{fig:lsllonefail} shows what would happen if the LL-LMS were to run in 
isolation. Starting at time 45, the LL-LMS sees that generator 2 has failed and it has 
switched to using generator 1 to power both buses according to its priority table. In order
to stay under generator 1's power rating, the LL-LMS sheds a large number of loads on the 
second bus. Figure~\ref{fig:lsolonefail} shows what would happen in a combined HL-LMS
LL-LMS system. The HL-LMS produces advice to be used from time 40 to 50. Unfortunately,
there is no way for this advice to be aware of the failure at time 45. So while the advice
from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. The LL-LMS is
able to see this at time 45 and kicks in, ignoring the unsafe advice from the HL-LMS. This
behavior is observed in Figure~\ref{fig:lsolonefail} where from time 45 to 49, a large
number of loads are shed. Then at time 50, the HL-LMS reoptimizes by turning on the APU.
Notice the difference in the number of loads shed onward from time 50 in 
Figure~\ref{fig:lsllonefail} and Figure~\ref{fig:lsolonefail}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/lsllonefail.png}
  \caption{\textbf{LL-LMS Load Shedding -- One Failure}. Notice
  the large amounts of load shedded in order to keep power usage below the power rating
  of generator 1.}
  \label{fig:lsllonefail}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
  \caption{\textbf{OL-LMS Load Shedding -- One Failure}. Notice the
  large number of loads dropped from time 45 to 50 before the HL-LMS can reoptimize.}
  \label{fig:lsolonefail}
\end{figure}

\subsection{Unexpected Power Spikes}
The HL-LMS performs a live optimization using historical data for load power requests from
previous flights. Due to the nature of the optimization problem, it is impossible to
program the optimization problem with live data. As a result, it is possible for the
HL-LMS to provide unsafe advice if there is a power spike. More specifically, it is possible
for the HL-LMS to generate a load shedding configuration that could have one of the power
generators driving a higher power request than its rating. We want to simulate such a
failure case and show that the LL-LMS performs the required adjustment to keep the system
in a safe state. For this experiment, we use the same historical power request as in
Figure~\ref{fig:preqnofail}. However, the actual power request for this simulation is shown
in Figure~\ref{fig:preqpwrspike}. Without the LL-LMS, the HL-LMS would use the load
shedding configuration seen in Figure~\ref{fig:lshlnofail}. Notice that there is a power
spike from time 10 to 40 that the historical data does not account for and thus the HL-LMS
is unaware of. As a result, its load shedding configuration would overload the power
generator. Because we have a LL-LMS, we are able to keep the system in a safe configuration.
Figure~\ref{fig:lsolpwrspike} shows that the LL-LMS will detect that the generators would
be overloaded and thus shed some loads during this time interval.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/preqpwrspike.png}
  \caption{\textbf{Actual Power Request} Notice the difference between this power
  request and the power request expected by the historical data in 
  Figure~\ref{fig:preqnofail}. In particular, there is a large power spike from time 10
  to 40 that is not accounted for in the historical data.}
  \label{fig:preqpwrspike}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
  \caption{\textbf{OL-LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
  use the advice from the HL-LMS. The only time it ignores the advice from the HL-LMS is
  from time 10 to 40 when there is a power spike that the HL-LMS is not aware of.}
  \label{fig:lsolpwrspike}
\end{figure}

\section{Related Work}
Related~\cite{Lamport:LaTeX}.

\section{Conclusion}
Conclusion.

\section*{Acknowledgments}
This work was supported by the US Department of Defense (DoD) through the National Defense Science \& Engineering Graduate Fellowship (NDSEG) Program.

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2012-12-02 project/debug_mehdi_code/OptProb_Linear.m
getting ready to run our checkSafety() on the 100% off-the-shelf code from Mehdi

@@ -1,0 +1,196 @@
+% Choose between different load files, load1, load2, load3,...
+
+clear all
+%% constants
+Nt=10+1;   % number of time steps   % can select: 10+1, 20+1, 50+1 and 100+1.
+Nl=10;   % number of loads connected to each bus
+Ns=3;    % number of power sources
+Nb=2;    % number of HVAC buses
+
+N=100;   % length of prediction horizon
+
+%% load the "loads"
+[Ls1,Lns1,Ls2,Lns2]=load3(N);   % choose between load1, load2 and load3.
+
+%% Plot the load requirements 
+
+%Load requirements for bus 1
+Ls1sum=sum(Ls1,1); 
+Lns1sum=sum(Lns1,1);
+
+figure;
+subplot(2,1,1);
+plot(1:1:N,Lns1sum,'--b','LineWidth',2);
+title('P_{req}(t) for bus 1');
+xlabel('time [s]');
+ylabel('P_{req_1} [W]');
+grid on;
+L1sum=Ls1sum+Lns1sum;
+hold on;
+plot(1:1:N,L1sum,'-r','LineWidth',2);
+legend('L_{ns}','L_{s}+L_{ns}');
+axis([1 N 0 1.1*max(L1sum)]);
+
+
+% Load requirements for bus 2
+Ls2sum=sum(Ls2,1); 
+Lns2sum=sum(Lns2,1); 
+
+%figure;
+subplot(2,1,2);
+plot(1:1:N,Lns2sum,'--b','LineWidth',2);
+title('P_{req}(t) for bus 2');
+xlabel('time [s]');
+ylabel('P_{req_2} [W]');
+grid on;
+L2sum=Ls2sum+Lns2sum;
+hold on;
+plot(1:1:N,L2sum,'-r','LineWidth',2);
+legend('L_{ns}','L_{s}+L_{ns}');
+axis([1 N 0 1.1*max(L2sum)]);
+
+%% Max. Power supply by Engines and APU
+
+U1=1e5;     Peng1=U1*ones(1,Nt);
+U2=1e5;     Peng2=U2*ones(1,Nt);
+U3=104e3;   Papu=U3*ones(1,Nt);
+U=[U1*ones(1,Nt); U2*ones(1,Nt); U3*ones(1,Nt)];
+P=[Peng1' Peng2' Papu'];
+
+%% Optimization problem set-up
+tic;
+%Coefficients:
[…]

### 2012-12-02 project/debug_mehdi_code/load2.m

@@ -1,0 +1,45 @@
+function [Ls1,Lns1,Ls2,Lns2]=load2(N)
+
+% BUS 1
+% Non-Sheddable loads
+
+
+Ls1(1,:)=-25000*ones(1,N);
+Ls1(2,:)=zeros(1,N); Ls1(2,N/10:end)=1000;
+Ls1(3,:)=zeros(1,N); Ls1(3,2*N/10:9*N/10)=1000;
+Ls1(4,:)=zeros(1,N); Ls1(4,N/10:9*N/10)=2000;
+Ls1(5,:)=zeros(1,N); Ls1(5, 2*N/10:9*N/10 )=1000;
+Ls1(6,:)=zeros(1,N); Ls1(6, 3*N/10:9*N/10 )=1000;
+Ls1(7,:)=zeros(1,N); Ls1(7, 2*N/10:5*N/10 )=5000; Ls1(7, 7*N/10:9*N/10 )=5000;
+Ls1(8,:)=zeros(1,N); Ls1(8, 3*N/10:5*N/10 )=5000; Ls1(8, 8*N/10:9*N/10 )=6000;
+Ls1(9,:)=zeros(1,N); Ls1(9, 4*N/10:9*N/10 )=500; 
+Ls1(10,:)=zeros(1,N); Ls1(10, 4*N/10:6*N/10 )=500;
+
+
+% Sheddable loads
+Lns1(1,:)=41000*ones(1,N);   Lns1(2,:)=5000*ones(1,N);     Lns1(3,:)=2000*ones(1,N);
+Lns1(4,:)=2000*ones(1,N);    Lns1(5,:)=1000*ones(1,N);     Lns1(6,:)=5000*ones(1,N);
+Lns1(7,:)=1000*ones(1,N);    Lns1(8,:)=2000*ones(1,N);     Lns1(9,:)=2000*ones(1,N);
+Lns1(10,:)=1000*ones(1,N);
+
+
+% BUS 2
+% Non-Sheddable loads
+Ls2(1,:)=-25000*ones(1,N);  % 25000
+Ls2(2,:)=zeros(1,N); Ls2(2,N/10:end)=1000;
+Ls2(3,:)=zeros(1,N); Ls2(3,1:9*N/10)=1000; 
+Ls2(4,:)=zeros(1,N); Ls2(4,1:9*N/10)=2000;
+Ls2(5,:)=zeros(1,N); Ls2( 5,2*N/10:9*N/10 )=1000;
[…]

### 2012-12-02 project/genWorkload.m

@@ -10,10 +10,10 @@ end
 
 %perturb one set of workloads on a bus (e.g. Ls1, Lns1, ...)
 function bus =  perturbBus(bus)
-    randRange = 5.2; %will use random numbers in the space (-randRange/2, randRange/2)
+    randRange = 4.0; %will use random numbers in the space (-randRange/2, randRange/2)
     Nl = size(bus); %number of individual loads to perturb
     perturbation = (rand(Nl) * (randRange)) - (randRange/2); %random vector of size Nl with the specified range, centered at 0
-    bus = abs(bus + bus .* perturbation)
+    bus = abs(bus + bus .* perturbation);
 
 end
 

### 2012-12-02 project/debug_mehdi_code/checkSafety.m
forking off a version of checkSafety that can interface directly to Mehdi's stuff

@@ -1,0 +1,46 @@
+
+% Are any power caps exceeded?
+function isSafe = checkSafety(config, sensors, constants)
+
+    Shedding1 = config.Shedding1;
+    Shedding2 = config.Shedding2;
+    Bus1_pwrReq = sum(sensors.workload.Ls1) + sum(sensors.workload.Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
+    Bus2_pwrReq = sum(sensors.workload.Ls2) + sum(sensors.workload.Lns2(Shedding2>0));
+   
+    %calculate pwr required from each generator
+    pwrReqGen1 = 0; pwrReqGen2 = 0; pwrReqApu = 0; 
+    if (config.BusGen(1) == 1)
+        pwrReqGen1 = Bus1_pwrReq;
+    elseif (config.BusGen(1) == 2)
+        pwrReqGen2 = Bus1_pwrReq;
+    elseif (config.BusGen(1) == 3)
+        pwrReqApu = Bus1_pwrReq;
+    end
+
+    if (config.BusGen(2) == 1)
+        pwrReqGen1 = pwrReqGen1 + Bus2_pwrReq;
+    elseif (config.BusGen(2) == 2)
+        pwrReqGen2 = pwrReqGen2 + Bus2_pwrReq;
+    elseif (config.BusGen(2) == 3)
+        pwrReqApu = pwrReqApu + Bus2_pwrReq;
+    end
+
+    %decide whether any of the generators are being overdrawn
+    isSafe=1;
+    
+    % check generator status
+    if (sensors.genStatus(config.BusGen(1)) == 0 || sensors.genStatus(config.BusGen(2)) == 0) %have we assigned a broken generator to a bus?
+        isSafe=0;
+    end
+    
+    % checking for generator overload
+    if(pwrReqGen1 > constants.generatorOutput(1))
+        isSafe=0;
+    elseif(pwrReqGen2 > constants.generatorOutput(2))
+        isSafe=0;
+    elseif(pwrReqApu > constants.generatorOutput(3))
+        isSafe=0;
+    end
+end
+
+

### 2012-12-02 project/debug_mehdi_code/checkSafety.m
revising (forked off version of) checkSafety for Mehdi's interface

@@ -1,27 +1,31 @@
 
 % Are any power caps exceeded?
-function isSafe = checkSafety(config, sensors, constants)
+%function isSafe = checkSafety(config, sensors, constants)
 
-    Shedding1 = config.Shedding1;
-    Shedding2 = config.Shedding2;
-    Bus1_pwrReq = sum(sensors.workload.Ls1) + sum(sensors.workload.Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
-    Bus2_pwrReq = sum(sensors.workload.Ls2) + sum(sensors.workload.Lns2(Shedding2>0));
+%note: all 'binvar' objects are cast to double.
+function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAssignment1, genAssignment2, generatorOutput) 
+
+    %TODO: convert genAssignment to BusGen
+    genStatus = [1 1 1]; %assume no generator failures for this experiment
+
+    Bus1_pwrReq = sum(Ls1) + sum(Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
+    Bus2_pwrReq = sum(Ls2) + sum(Lns2(Shedding2>0));
    
     %calculate pwr required from each generator
     pwrReqGen1 = 0; pwrReqGen2 = 0; pwrReqApu = 0; 
-    if (config.BusGen(1) == 1)
+    if (BusGen(1) == 1)
         pwrReqGen1 = Bus1_pwrReq;
-    elseif (config.BusGen(1) == 2)
+    elseif (BusGen(1) == 2)
         pwrReqGen2 = Bus1_pwrReq;
-    elseif (config.BusGen(1) == 3)
+    elseif (BusGen(1) == 3)
         pwrReqApu = Bus1_pwrReq;
     end
 
-    if (config.BusGen(2) == 1)
+    if (BusGen(2) == 1)
         pwrReqGen1 = pwrReqGen1 + Bus2_pwrReq;
-    elseif (config.BusGen(2) == 2)
+    elseif (BusGen(2) == 2)
         pwrReqGen2 = pwrReqGen2 + Bus2_pwrReq;
-    elseif (config.BusGen(2) == 3)
+    elseif (BusGen(2) == 3)
         pwrReqApu = pwrReqApu + Bus2_pwrReq;
     end
 
@@ -29,16 +33,16 @@ function isSafe = checkSafety(config, sensors, constants)
     isSafe=1;
     
     % check generator status
[…]

### 2012-12-02 project/debug_mehdi_code/OptProb_Linear.m
reproduced the unsafe behavior with off-the-shelf code from Mehdi

@@ -1,8 +1,9 @@
 % Choose between different load files, load1, load2, load3,...
 
+close all hidden %get rid of old figures
 clear all
 %% constants
-Nt=10+1;   % number of time steps   % can select: 10+1, 20+1, 50+1 and 100+1.
+Nt=100+1;   % number of time steps   % can select: 10+1, 20+1, 50+1 and 100+1.
 Nl=10;   % number of loads connected to each bus
 Ns=3;    % number of power sources
 Nb=2;    % number of HVAC buses
@@ -110,6 +111,13 @@ obj = obj + M * sum(sum(alpha));
 options=sdpsettings('solver','cplex');
 solvesdp(cons,obj,options);
 toc;
+
+generatorOutput = [U1 U2 U3]; %how much pwr each generator can produce
+isSafe = zeros([1 N]); 
+for i=1:N
+    isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput); 
+end
+display(isSafe)
 
 % Plot C
 xp=1:1:Nt*100/(Nt-1);  % 110

### 2012-12-02 project/debug_mehdi_code/checkSafety.m

@@ -5,7 +5,10 @@
 %note: all 'binvar' objects are cast to double.
 function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAssignment1, genAssignment2, generatorOutput) 
 
-    %TODO: convert genAssignment to BusGen
+    BusGen = [0 0];
+    [myMax BusGen(1)] = max(genAssignment1);  %BusGen(1) is argmax here
+    [myMax BusGen(2)] = max(genAssignment2);
+
     genStatus = [1 1 1]; %assume no generator failures for this experiment
 
     Bus1_pwrReq = sum(Ls1) + sum(Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate

### 2012-12-02 project/debug_mehdi_code/OptProb_Linear.m
packing up a nice piece of code to show Mehdi, for reproducing the 'HL-LMS unsafe' stuff

@@ -115,9 +115,9 @@ toc;
 generatorOutput = [U1 U2 U3]; %how much pwr each generator can produce
 isSafe = zeros([1 N]); 
 for i=1:N
-    isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput); 
+    isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput, i); 
 end
-display(isSafe)
+%display(isSafe)
 
 % Plot C
 xp=1:1:Nt*100/(Nt-1);  % 110

### 2012-12-02 project/debug_mehdi_code/checkSafety.m

@@ -3,7 +3,7 @@
 %function isSafe = checkSafety(config, sensors, constants)
 
 %note: all 'binvar' objects are cast to double.
-function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAssignment1, genAssignment2, generatorOutput) 
+function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAssignment1, genAssignment2, generatorOutput, time) 
 
     BusGen = [0 0];
     [myMax BusGen(1)] = max(genAssignment1);  %BusGen(1) is argmax here
@@ -43,9 +43,12 @@ function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAss
     % checking for generator overload
     if(pwrReqGen1 > generatorOutput(1))
         isSafe=0;
+        sprintf('UNSAFE at time %d: Generator 1 produces %d power units, but the HL-LMS allocated Generator 1 to provide %d power units', time, generatorOutput(1), pwrReqGen1)
     elseif(pwrReqGen2 > generatorOutput(2))
         isSafe=0;
+        sprintf('UNSAFE at time %d: Generator 2 produces %d power units, but the HL-LMS allocated Generator 2 to provide %d power units', time, generatorOutput(2), pwrReqGen2)
     elseif(pwrReqApu > generatorOutput(3))
+        sprintf('UNSAFE at time %d: APU produces %d power units, but the HL-LMS allocated APU to provide %d power units', time, generatorOutput(3), pwrReqApu)
         isSafe=0;
     end
 end

### 2012-12-02 project/checkSafety.m
sent reproducable explanation of the problem to Mehdi

@@ -3,7 +3,7 @@
 function isSafe = checkSafety(config, sensors, constants)
     Shedding1 = config.Shedding1;
     Shedding2 = config.Shedding2;
-    Bus1_pwrReq = sum(sensors.workload.Ls1) + sum(sensors.workload.Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
+    Bus1_pwrReq = sum(sensors.workload.Ls1) + sum(sensors.workload.Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate'
     Bus2_pwrReq = sum(sensors.workload.Ls2) + sum(sensors.workload.Lns2(Shedding2>0));
    
     %calculate pwr required from each generator

### 2012-12-02 project/plotGraphs.m
playing with line width in matlab plots

@@ -69,6 +69,7 @@ function plotC(C1, C2, Nt, N, xp)
     set(gca,'YTick',0:1:1);
     set(gca,'YTickLabel',{'Shed (off)','Granted (on)'});
     xlabel('time [s]');
+    set(gca, 'LineWidth', 5.0) 
 
     subplot(2,2,2);
     plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:))

### 2012-12-02 project/plotGraphs.m
linewidth=3.0 is kinda too thick

@@ -61,7 +61,7 @@ function plotC(C1, C2, Nt, N, xp)
  
     figure;
     subplot(2,2,1);
-    plot(xp,C1(1,:),xp,C1(2,:),xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',2)
+    plot(xp,C1(1,:),xp,C1(2,:),xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',3.0)
 
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 1 to 5)');
@@ -69,10 +69,9 @@ function plotC(C1, C2, Nt, N, xp)
     set(gca,'YTick',0:1:1);
     set(gca,'YTickLabel',{'Shed (off)','Granted (on)'});
     xlabel('time [s]');
-    set(gca, 'LineWidth', 5.0) 
 
     subplot(2,2,2);
-    plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:))
+    plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:), 'LineWidth',3.0)
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 2 (loads 1 to 5)');
     axis([0 N+10 -0.1 1.5]);
@@ -81,7 +80,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,3);
-    plot(xp,C1(6,:),xp,C1(7,:),xp,C1(8,:),xp,C1(9,:),xp,C1(10,:))
+    plot(xp,C1(6,:),xp,C1(7,:),xp,C1(8,:),xp,C1(9,:),xp,C1(10,:), 'LineWidth',3.0)
     legend('L_6','L_7','L_8','L_9','L_{10}','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 6 to 10)');
     axis([0 N+10 -0.1 1.5]);
@@ -90,7 +89,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,4);
[…]

### 2012-12-02 project/plotGraphs.m
eh, the wider graph lines make the plots look really 'choppy' -- let's forget the wide lines thing for now

@@ -61,7 +61,7 @@ function plotC(C1, C2, Nt, N, xp)
  
     figure;
     subplot(2,2,1);
-    plot(xp,C1(1,:),xp,C1(2,:),xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',3.0)
+    plot(xp,C1(1,:),xp,C1(2,:),xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',1.5)
 
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 1 to 5)');
@@ -71,7 +71,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,2);
-    plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:), 'LineWidth',3.0)
+    plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:), 'LineWidth',1.5)
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 2 (loads 1 to 5)');
     axis([0 N+10 -0.1 1.5]);
@@ -80,7 +80,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,3);
-    plot(xp,C1(6,:),xp,C1(7,:),xp,C1(8,:),xp,C1(9,:),xp,C1(10,:), 'LineWidth',3.0)
+    plot(xp,C1(6,:),xp,C1(7,:),xp,C1(8,:),xp,C1(9,:),xp,C1(10,:), 'LineWidth',1.5)
     legend('L_6','L_7','L_8','L_9','L_{10}','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 6 to 10)');
     axis([0 N+10 -0.1 1.5]);
@@ -89,7 +89,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,4);
-    plot(xp,C2(6,:),xp,C2(7,:),xp,C2(8,:),xp,C2(9,:),xp,C2(10,:), 'LineWidth',3.0)
[…]

### 2012-12-02 project/debug_mehdi_code/checkSafety.m
implementing stuff from Mehdi's email (is OLMS indeed going over power budget?

@@ -11,8 +11,8 @@ function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAss
 
     genStatus = [1 1 1]; %assume no generator failures for this experiment
 
-    Bus1_pwrReq = sum(Ls1) + sum(Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
-    Bus2_pwrReq = sum(Ls2) + sum(Lns2(Shedding2>0));
+    Bus1_pwrReq = sum(Lns1) + sum(Ls1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
+    Bus2_pwrReq = sum(Lns2) + sum(Ls2(Shedding2>0));
    
     %calculate pwr required from each generator
     pwrReqGen1 = 0; pwrReqGen2 = 0; pwrReqApu = 0; 

### 2012-12-03 project/latex_for_presentation/test.tex
latex for the one equation we are putting in our ppt slides

\documentclass[10pt,journal,letterpaper,compsoc, draftclsnofoot, onecolumn]{IEEEtran} %draft mode for double-spaced
\usepackage{graphicx}
\usepackage{subfig} % for horizontal figures
\usepackage{amsmath}
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
\usepackage{cite}

\begin{document}

% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Title Here}

\author{Forrest~N.~Iandola,~\IEEEmembership{Student Member,~IEEE,}
        Praveen~Jayachandran,~\IEEEmembership{Member,~IEEE,}
        and~Tarek~Abdelzaher,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem Forrest N. Iandola is with the Department of Computer Science, University of Illinois at Urbana-Champaign, Urbana, IL, 61801.
\protect\\
E-mail: iandola1@illinois.edu
\IEEEcompsocthanksitem Praveen Jayachandran is with IBM Research, India.
\protect\\
E-mail: prjayach@in.ibm.com
\IEEEcompsocthanksitem Tarek Abdelzaher is with the Department of Computer Science, University of Illinois at Urbana-Champaign, Urbana, IL, 61801.
\protect\\
E-mail: zaher@illinois.edu}% <-this % stops a space
\thanks{}}

% The paper headers
\markboth{IEEE Transactions on Computers}%
{Iandola, Jayahandran, and Abdelzaher: Optimizing the Structural Robustness of Real-Time Distributed Systems Towards Uncertainties in Execution Time}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
Hello I'm an abstract.

\end{abstract}

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
\noindent
C.3 [Special-purpose and Application-based Systems]: Real-time and embedded systems \\
C.2.4 [Computer-communication Networks]: Distributed Systems \\
C.2.2 [Computer-communication Networks]: Network Protocols -- Routing protocols \\
C.2.8 [Mobile Computing]: Algorithm/protocol design and analysis \\

\noindent
Robustness, Real-Time Systems, Distributed Systems, Networks, Delay Uncertainties, Execution Overruns.
\end{keywords}}

\maketitle
\IEEEdisplaynotcompsoctitleabstractindextext
\IEEEpeerreviewmaketitle

\section{Introduction}

\begin{figure}[htb]
  \centering
  \fbox{
      \subfloat[]{\label{fig:delayAmtRandomized_proportionalToPriority}  TODO }    
  }
  \fbox{            
      \subfloat[]{\label{fig:delayAmtRandomized_random} TODO}
  }
  \caption{Influence of extent to which tasks are delayed on robustness, using randomized WCET overrun percentages. Left: importance equal to priority; Right: random importance values.}
  \label{fig:delayAmtRandomized_doubleFigure}
\end{figure}

I'm trying to reference the ``main" number (not $a$ or $b$) of a subfloat figure.
Figure~\ref{fig:delayAmtRandomized_doubleFigure}.
AHA! The trick was to put label{} after caption{} in the figure.

\vspace{-.5in}
\begin{IEEEbiographynophoto}{Forrest Iandola}
earned the BS degree in Computer Science at University of Illinois at Urbana-Champaign in 2012, and he is now pursuing his PhD in Computer Science at University of California, Berkeley. 
His BS thesis focused on scheduling and routing methodologies for real-time distributed systems. 
His PhD research, funded by the National Defense Science and Engineering Graduate (NDSEG) Fellowship, focuses on designing fast and scalable computer vision algorithms for real-time sensing and interactive systems. 
Forrest is also interested in scientific computing, and he designed large-scale physics simulations at three US Department of Energy National Laboratories.
He is the recipient of the Daniel L. Slotnick Scholarship from the UIUC Department of Computer Science, and he is a member of IEEE, SIAM, and ACM.
\end{IEEEbiographynophoto}
\vspace{-.7in}
\begin{IEEEbiographynophoto}{Praveen Jayachandran}
is a research staff member of the next generation
services and cloud computing team at IBM Research, India. His research
interests are in the area of analysis, performance management, and
optimization of distributed systems and networks. He received his PhD from
the University of Illinois at Urbana-Champaign in 2010. He is the recipient
of the best paper award at ECRTS 2009, the best student paper award at
ECRTS 2007, and the C.L. and Jane Liu award from the Department of Computer
Science at UIUC in 2007.
\end{IEEEbiographynophoto}
\vspace{-.7in}

\begin{IEEEbiographynophoto}{Tarek Abdelzaher}
%Tarek's bio from "Efficient flow-control algorithm cooperating with energy allocation scheme for solar-powered WSNs" paper
received his B.Sc. and M.Sc. degrees in Electrical and Computer Engineering from Ain Shams University, Cairo, Egypt, in 1990 and 1994 respectively. 
He received his Ph.D. from the University of Michigan in 1999 on Quality of Service Adaptation in Real-Time Systems. 
He has been an Assistant Professor at the University of Virginia, where he founded the Software Predictability Group until 2005. 
He is currently a Full Professor at the Department of Computer Science, the University of Illinois at Urbana Champaign. 
He has authored/coauthored more than 150 refereed publications in real-time computing, distributed systems, sensor networks, and control. 
He is Editor-in-Chief of the Journal of Real-Time Systems, an Associate Editor of the IEEE Transactions on Mobile Computing, IEEE Transactions on Parallel and Distributed Systems, the ACM Transaction on Sensor Networks, and the Ad Hoc Networks Journal. 
%He was Program Chair of RTAS 2004, RTSS 2006, IPSN 2010 and ICDCS 2010, as well as General Chair of RTAS 2005, IPSN 2007, RTSS 2007, DCoSS 2008, and Sensys 2008. 
Abdelzaher's research interests lie broadly in understanding and controlling performance and temporal properties of networked embedded and software systems in the face of increasing complexity, distribution, and degree of embedding in an external physical environment. 
Tarek Abdelzaher is a member of IEEE and ACM.
\end{IEEEbiographynophoto}

\end{document}

### 2012-12-03 project/latex_for_presentation/test.tex
latex stuff

@@
 
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Title Here}
-
-\author{Forrest~N.~Iandola,~\IEEEmembership{Student Member,~IEEE,}
-        Praveen~Jayachandran,~\IEEEmembership{Member,~IEEE,}
-        and~Tarek~Abdelzaher,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
-\IEEEcompsocitemizethanks{
-\IEEEcompsocthanksitem Forrest N. Iandola is with the Department of Computer Science, University of Illinois at Urbana-Champaign, Urbana, IL, 61801.
-\protect\\
-E-mail: iandola1@illinois.edu
-\IEEEcompsocthanksitem Praveen Jayachandran is with IBM Research, India.
-\protect\\
-E-mail: prjayach@in.ibm.com
-\IEEEcompsocthanksitem Tarek Abdelzaher is with the Department of Computer Science, University of Illinois at Urbana-Champaign, Urbana, IL, 61801.
-\protect\\
-E-mail: zaher@illinois.edu}% <-this % stops a space
-\thanks{}}
-
-% The paper headers
-\markboth{IEEE Transactions on Computers}%
-{Iandola, Jayahandran, and Abdelzaher: Optimizing the Structural Robustness of Real-Time Distributed Systems Towards Uncertainties in Execution Time}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-
-% for Computer Society papers, we must declare the abstract and index terms
-% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
-% command as these need to go into the title area created by \maketitle.
-\IEEEcompsoctitleabstractindextext{%
-\begin{abstract}
-Hello I'm an abstract.
-
-\end{abstract}
-
-% Note that keywords are not normally used for peer review papers.
-\begin{keywords}
-\noindent
-C.3 [Special-purpose and Application-based Systems]: Real-time and embedded systems \\
-C.2.4 [Computer-communication Networks]: Distributed Systems \\
-C.2.2 [Computer-communication Networks]: Network Protocols -- Routing protocols \\
-C.2.8 [Mobile Computing]: Algorithm/protocol design and analysis \\
-
-\noindent
-Robustness, Real-Time Systems, Distributed Systems, Networks, Delay Uncertainties, Execution Overruns.
-\end{keywords}}
-
-\maketitle
-\IEEEdisplaynotcompsoctitleabstractindextext
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-
-\begin{figure}[htb]
-  \centering
-  \fbox{
-      \subfloat[]{\label{fig:delayAmtRandomized_proportionalToPriority}  TODO }    
-  }
-  \fbox{            
-      \subfloat[]{\label{fig:delayAmtRandomized_random} TODO}
-  }
-  \caption{Influence of extent to which tasks are delayed on robustness, using randomized WCET overrun percentages. Left: importance equal to priority; Right: random importance values.}
-  \label{fig:delayAmtRandomized_doubleFigure}
-\end{figure}
-
-I'm trying to reference the ``main" number (not $a$ or $b$) of a subfloat figure.
-Figure~\ref{fig:delayAmtRandomized_doubleFigure}.
-AHA! The trick was to put label{} after caption{} in the figure.
-
-\vspace{-.5in}
-\begin{IEEEbiographynophoto}{Forrest Iandola}
-earned the BS degree in Computer Science at University of Illinois at Urbana-Champaign in 2012, and he is now pursuing his PhD in Computer Science at University of California, Berkeley. 
-His BS thesis focused on scheduling and routing methodologies for real-time distributed systems. 
-His PhD research, funded by the National Defense Science and Engineering Graduate (NDSEG) Fellowship, focuses on designing fast and scalable computer vision algorithms for real-time sensing and interactive systems.
-Forrest is also interested in scientific computing, and he designed large-scale physics simulations at three US Department of Energy National Laboratories.
-He is the recipient of the Daniel L. Slotnick Scholarship from the UIUC Department of Computer Science, and he is a member of IEEE, SIAM, and ACM.
-\end{IEEEbiographynophoto}
-\vspace{-.7in}
-\begin{IEEEbiographynophoto}{Praveen Jayachandran}
-is a research staff member of the next generation
-services and cloud computing team at IBM Research, India. His research
-interests are in the area of analysis, performance management, and
-optimization of distributed systems and networks. He received his PhD from
-the University of Illinois at Urbana-Champaign in 2010. He is the recipient
-of the best paper award at ECRTS 2009, the best student paper award at
-ECRTS 2007, and the C.L. and Jane Liu award from the Department of Computer
-Science at UIUC in 2007.
-\end{IEEEbiographynophoto}
-\vspace{-.7in}
-
-\begin{IEEEbiographynophoto}{Tarek Abdelzaher}
-%Tarek's bio from "Efficient flow-control algorithm cooperating with energy allocation scheme for solar-powered WSNs" paper
-received his B.Sc. and M.Sc. degrees in Electrical and Computer Engineering from Ain Shams University, Cairo, Egypt, in 1990 and 1994 respectively. 
-He received his Ph.D. from the University of Michigan in 1999 on Quality of Service Adaptation in Real-Time Systems. 
-He has been an Assistant Professor at the University of Virginia, where he founded the Software Predictability Group until 2005. 
-He is currently a Full Professor at the Department of Computer Science, the University of Illinois at Urbana Champaign. 
-He has authored/coauthored more than 150 refereed publications in real-time computing, distributed systems, sensor networks, and control. 
-He is Editor-in-Chief of the Journal of Real-Time Systems, an Associate Editor of the IEEE Transactions on Mobile Computing, IEEE Transactions on Parallel and Distributed Systems, the ACM Transaction on Sensor Networks, and the Ad Hoc Networks Journal.
-%He was Program Chair of RTAS 2004, RTSS 2006, IPSN 2010 and ICDCS 2010, as well as General Chair of RTAS 2005, IPSN 2007, RTSS 2007, DCoSS 2008, and Sensys 2008. 
-Abdelzaher's research interests lie broadly in understanding and controlling performance and temporal properties of networked embedded and software systems in the face of increasing complexity, distribution, and degree of embedding in an external physical environment.
-Tarek Abdelzaher is a member of IEEE and ACM.
-\end{IEEEbiographynophoto}
+blah blah i'm some math
 

### 2012-12-03 project/debug_mehdi_code/OptProb_Linear.m
about to test an idea for an off-by-one error in data from Mehdi's code

@@ -115,7 +115,8 @@ toc;
 generatorOutput = [U1 U2 U3]; %how much pwr each generator can produce
 isSafe = zeros([1 N]); 
 for i=1:N
-    isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput, i); 
+    %isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput, i); 
+    isSafe(i) = checkSafety(double(C1(:,i+1)), double(C2(:,i+1)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i+1)), double(Del2(:,+1)), generatorOutput, i); 
 end
 %display(isSafe)
 

### 2012-12-03 project/debug_mehdi_code/checkSafety.m

@@ -11,8 +11,8 @@ function isSafe = checkSafety(Shedding1, Shedding2, Ls1, Lns1, Ls2, Lns2, genAss
 
     genStatus = [1 1 1]; %assume no generator failures for this experiment
 
-    Bus1_pwrReq = sum(Lns1) + sum(Ls1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate
-    Bus2_pwrReq = sum(Lns2) + sum(Ls2(Shedding2>0));
+    Bus1_pwrReq = sum(Lns1) + sum(Ls1(Shedding1<1)); %'Shedding1>0' is a Matlab 'logical coordinate
+    Bus2_pwrReq = sum(Lns2) + sum(Ls2(Shedding2<1));
    
     %calculate pwr required from each generator
     pwrReqGen1 = 0; pwrReqGen2 = 0; pwrReqApu = 0; 

### 2012-12-06 project/report/paper.tex
added content to final report

@@
 Overview, stuff about platform based design...
+
+Intro to aircraft EPS... explain the buses and generators.
+Possibly show the diagram from the Honeywell patent.
+Possibly show our ``cartoon" diagram of an EPS system.
+
+Our key objective is to design a Load Management System (LMS) for EPS that is ``smarter" in terms of several criteria.
+First, always preserve safety.
+After preserving safety, maximize the quality of service.
+Finally, conserve fuel by using as few generators as possible.
+
+Aircraft EPS generally have two types of loads: sheddable and non-sheddable.
+Examples of non-sheddable loads include navigation systems, electronic actuators, and other mission-critical devices.
+Sheddable loads include the cabin lights, kitchen, and electrical outlets for passengers' laptop computers.
+If an EPS power bus requires more power than the generators can provide, the LMS is responsible for disabling (shedding) some of the sheddable loads.
+
 \subsection{OL-LMS Goals}
 Talk about goals of project
-\section{HL-LMS}
-Reference Mehdi's et al work on HL-LMS. Brief description to show that we understand it.
-\section{LL-LMS}
-Talk about our LL-LMS system, its inputs/outputs, functionality, etc
+
+\section{Related Work}
+\label{sec:related-work}
+In typical aircraft EPS, the load management systems aim to ensure safety during unexpected system behavior [citation needed].
+Specifically, [these systems typically require] unsheddable loads to remain powered at all times. [or, maybe unsheddable loads can be unpowered for up to 50ms while switching generators]
+Therefore, if a generator fails, the EPS must shed loads and/or reallocate generators to buses such that no unsheddable load becomes unpowered.
+Typically, these systems rely on hard-coded priority tables for selecting loads to shed and for allocating generators to buses.
+
+\subsection{Optimizing-LMS}
+\label{sec:optimizing-LMS}
+LMS systems that are based purely on priority tables are effective for preserving safety.
+However, priority table solutions tend to shed more loads than necessary, to use more generators than necessary, [and be ill-defined in their tradeoffs between generator allocation and load shedding].
+Toward resolving these issues, the control systems community recently designed a model-predictive optimizing controller for aircraft EPS [Mehdi's paper].
+The key insight in [Mehdi's paper] is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
+Then, [Mehdi's paper] selected an objective function such that solving this optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
+[This optimizing controller aims to minimize three weighted parameters: load shedding preference, generator selection preference, and number of generators used.]
+[TODO: show a simplified version of Mehdi's math]
+
+However, the Optimizing LMS [Mehdi's paper] is not without shortcomings.
+Chiefly, the Optimizing LMS is computationally expensive (0.5s for 10 timesteps of optimization in CPLEX on an Intel 3930K processor).
+[Discuss receding horizon and the need for historical data]
+Since the Optimizing LMS relies on historical data, it may produce unsafe allocations that use a broken generator or operate above the generator power caps.
+Thus, it is potentially unsafe to use Optimizing LMS to control the electronics in a real airplane. 
+However, as we will see later in the paper, the Optimizing LMS can be effective if it is coupled with a fail-safe Low Level LMS.
+
+\section{Low Level LMS}
+[Talk about our LL-LMS system, its inputs/outputs, functionality, etc]
+
+This section presents a Low-Level LMS (LL-LMS) similar to the controllers discussed in Section~\ref{sec:related-work}.
+The key goal (and perhaps the only) goal of this load management system is to preserve safety if workloads change or if  generators fail.
+Therefore, using a pre-defined priority table of generator assignments and loads to shed, our LL-LMS sheds loads until the power requirement is within the generator's power budget.
+
+[TODO: 4-line pseudo-code for how we apply priority tables]
+
+Other details:
+Low-level controller samples system at every  timestep (e.g. 50ms of system time) 
+Ensure that no non-sheddable load is unpowered for more  than 50ms
+
+\section{Hierarchical LMS}
+Modify LL-LMS to call the Optimizing LMS periodically.
+LL-LMS does safety checks on results from Optimizing LMS...
+
 \section{OL-LMS}

### 2012-12-06 project/test_optimize_every_timestep/plotGraphs.m
bka

@@ -1,0 +1,136 @@
+
+% Expects ordinary arrays, not binvar arrays. (so, we got rid of '' and stuff)
+function plotGraphs(configLog, sensorLog, constants, Nt, N)
+
+    xp=1:1:Nt*100/(Nt-1);  % 110
+    plotPowerReq(constants.historicalWorkloads.Ls1(:,1:N), constants.historicalWorkloads.Lns1(:,1:N), constants.historicalWorkloads.Ls2(:,1:N), constants.historicalWorkloads.Lns2(:,1:N), N)
+
+    %Shedding1 = configLog(:).Shedding1; doesn't work -- it just returns the first timestep instead of all timesteps
+    Shedding1 = []; Shedding2 = [];
+    for i=1:N
+        Shedding1 = [Shedding1; configLog(i).Shedding1];
+        Shedding2 = [Shedding2; configLog(i).Shedding2];
+    end
+    plotC(Shedding1', Shedding2', Nt, N, xp)
+
+    %plotDelta(Del1, Del2, Nt, N, xp)
+    %plotBeta(Beta1, Beta2, Nt, N, xp)
+end
+
+function plotPowerReq(Ls1, Lns1, Ls2, Lns2, N)
+    %Load requirements for bus 1
+    Ls1sum=sum(Ls1,1); 
+    Lns1sum=sum(Lns1,1);
+
+    figure;
+    subplot(2,1,1);
+    plot(1:1:N,Lns1sum,'--b','LineWidth',2);
+    title('P_{req}(t) for bus 1');
+    xlabel('time [s]');
+    ylabel('P_{req_1} [W]');
+    grid on;
+    L1sum=Ls1sum+Lns1sum;
+    hold on;
+    plot(1:1:N,L1sum,'-r','LineWidth',2);
+    legend('L_{ns}','L_{s}+L_{ns}');
+    axis([1 N 0 1.1*max(L1sum)]);
+
+    % Load requirements for bus 2
+    Ls2sum=sum(Ls2,1); 
+    Lns2sum=sum(Lns2,1); 
+
+    %figure;
+    subplot(2,1,2);
+    plot(1:1:N,Lns2sum,'--b','LineWidth',2);
+    title('P_{req}(t) for bus 2');
+    xlabel('time [s]');
+    ylabel('P_{req_2} [W]');
+    grid on;
+    L2sum=Ls2sum+Lns2sum;
+    hold on;
[…]

### 2012-12-06 project/test_optimize_every_timestep/runAirplane.m

@@ -1,0 +1,46 @@
+%This is the "main function" for our airplane power simulation system
+function [] = runAirplane()
+    close all hidden %get rid of old figures
+
+    nTimesteps=100; % total number of timesteps in runAirplane outer loop 
+    Nl=10;   % number of loads connected to each bus. (10 sheddable, 10 unsheddable)
+    Ns=3;    % number of power sources
+    Nb=2;    % number of buses
+
+    %N and Nt are params for for Mehdi's code
+    N = 10; % prediction horizon
+    Nt = 10+1; % (prediction horizon + 1) -- some off-by-one-fix relic.
+
+    sensorLog = [];
+    configLog = [];
+    generatorOutput = [1e5, 1e5, 104e3]; %Pwr produced by generators. Called U1, U2, U3 in Mehdi's code
+        %TODO: make generatorOutput be a parameter to Mehdi's code, so that we can tweak it easily.
+    %[Ls1,Lns1,Ls2,Lns2]=load3(typecast(nTimesteps+Nt, 'int32')); % load the "loads" -- choose between load1, load2 and load3. (including extra padding, Nt, to avoid out-of-bounds in HLLMS)
+    [Ls1,Lns1,Ls2,Lns2]=load4(110);
+    historicalWorkloads = struct('Ls1', Ls1, 'Lns1', Lns1, 'Ls2', Ls2, 'Lns2', Lns2);
+    priorityTables = getPriorityTables();
+    constants = struct('historicalWorkloads', historicalWorkloads, 'priorityTables', priorityTables, 'generatorOutput', generatorOutput, 'nTimesteps', nTimesteps, 'Nt', Nt, 'Nl', Nl, 'Ns', Ns, 'Nb', Nb, 'N', N); %hard-coded params to pass around  
+    
+    advice = [];
+    HLclock = 1; %count up to each time we call the HLLMS
+    for time=1:nTimesteps
+        workload = genWorkload(historicalWorkloads, time);
+        genStatus = getGeneratorStatus(time);
+        sensors = struct('workload', workload, 'genStatus', genStatus, 'time', time);
+
+        if (HLclock == 1) %time to call HLLMS again
+            advice = HLLMS(sensors, constants);
+        end
+        config = LLLMS(sensors, constants, advice(HLclock));
+        if (HLclock == 10) HLclock = 1; 
+        else HLclock = HLclock + 1; 
+        end
+
[…]

### 2012-12-06 project/debug_mehdi_code/OptProb_Linear.m
test: doing horizon of 2 instead of horizon of 10. 0.35s per optimization (vs. 0.50s for horizon of 10)

@@ -116,7 +116,7 @@ generatorOutput = [U1 U2 U3]; %how much pwr each generator can produce
 isSafe = zeros([1 N]); 
 for i=1:N
     %isSafe(i) = checkSafety(double(C1(:,i)), double(C2(:,i)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i)), double(Del2(:,i)), generatorOutput, i); 
-    isSafe(i) = checkSafety(double(C1(:,i+1)), double(C2(:,i+1)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i+1)), double(Del2(:,+1)), generatorOutput, i); 
+    isSafe(i) = checkSafety(double(C1(:,i+1)), double(C2(:,i+1)), Ls1(:,i), Lns1(:,i), Ls2(:,i), Lns2(:,i), double(Del1(:,i+1)), double(Del2(:,i+1)), generatorOutput, i); 
 end
 %display(isSafe)
 

### 2012-12-06 project/test_optimize_every_timestep/runAirplane.m

@@ -8,8 +8,8 @@ function [] = runAirplane()
     Nb=2;    % number of buses
 
     %N and Nt are params for for Mehdi's code
-    N = 10; % prediction horizon
-    Nt = 10+1; % (prediction horizon + 1) -- some off-by-one-fix relic.
+    N = 2; % prediction horizon
+    Nt = 2+1; % (prediction horizon + 1) -- some off-by-one-fix relic.
 
     sensorLog = [];
     configLog = [];
@@ -28,13 +28,8 @@ function [] = runAirplane()
         genStatus = getGeneratorStatus(time);
         sensors = struct('workload', workload, 'genStatus', genStatus, 'time', time);
 
-        if (HLclock == 1) %time to call HLLMS again
-            advice = HLLMS(sensors, constants);
-        end
+        advice = HLLMS(sensors, constants);
         config = LLLMS(sensors, constants, advice(HLclock));
-        if (HLclock == 10) HLclock = 1; 
-        else HLclock = HLclock + 1; 
-        end
 
         configLog = [configLog config]; %this concatenation is slow ... but that's fine. 
         sensorLog = [sensorLog sensors];

### 2012-12-06 project/plotGraphs.m
modifying plots so that all the lines don't overlap

@@ -61,7 +61,7 @@ function plotC(C1, C2, Nt, N, xp)
  
     figure;
     subplot(2,2,1);
-    plot(xp,C1(1,:),xp,C1(2,:),xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',1.5)
+    plot(xp,C1(1,:),xp,C1(2,:)+0.01,xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',1.5)
 
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 1 to 5)');

### 2012-12-06 project/plotGraphs.m
now, shedding graph has stratified lines so that the lines dont overlap

@@ -61,7 +61,7 @@ function plotC(C1, C2, Nt, N, xp)
  
     figure;
     subplot(2,2,1);
-    plot(xp,C1(1,:),xp,C1(2,:)+0.01,xp,C1(3,:),xp,C1(4,:),xp,C1(5,:), 'LineWidth',1.5)
+    plot(xp,C1(1,:),xp,C1(2,:)+0.02,xp,C1(3,:)+0.04,xp,C1(4,:)+0.06,xp,C1(5,:)+0.08, 'LineWidth',1.5)
 
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 1 to 5)');
@@ -71,7 +71,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,2);
-    plot(xp,C2(1,:),xp,C2(2,:),xp,C2(3,:),xp,C2(4,:),xp,C2(5,:), 'LineWidth',1.5)
+    plot(xp,C2(1,:),xp,C2(2,:)+0.02,xp,C2(3,:)+0.04,xp,C2(4,:)+0.06,xp,C2(5,:)+0.08, 'LineWidth',1.5)
     legend('L_1','L_2','L_3','L_4','L_5','Orientation','horizontal');
     title('Power shedding of AC bus 2 (loads 1 to 5)');
     axis([0 N+10 -0.1 1.5]);
@@ -80,7 +80,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,3);
-    plot(xp,C1(6,:),xp,C1(7,:),xp,C1(8,:),xp,C1(9,:),xp,C1(10,:), 'LineWidth',1.5)
+    plot(xp,C1(6,:),xp,C1(7,:)+0.02,xp,C1(8,:)+0.04,xp,C1(9,:)+0.06,xp,C1(10,:)+0.08, 'LineWidth',1.5)
     legend('L_6','L_7','L_8','L_9','L_{10}','Orientation','horizontal');
     title('Power shedding of AC bus 1 (loads 6 to 10)');
     axis([0 N+10 -0.1 1.5]);
@@ -89,7 +89,7 @@ function plotC(C1, C2, Nt, N, xp)
     xlabel('time [s]');
 
     subplot(2,2,4);
[…]

### 2012-12-06 project/runAirplane.m
genWorkload takes additional parameter to decide whether or not to alter workload

@@ -16,7 +16,7 @@ function [] = runAirplane()
     generatorOutput = [1e5, 1e5, 104e3]; %Pwr produced by generators. Called U1, U2, U3 in Mehdi's code
         %TODO: make generatorOutput be a parameter to Mehdi's code, so that we can tweak it easily.
     %[Ls1,Lns1,Ls2,Lns2]=load3(typecast(nTimesteps+Nt, 'int32')); % load the "loads" -- choose between load1, load2 and load3. (including extra padding, Nt, to avoid out-of-bounds in HLLMS)
-    [Ls1,Lns1,Ls2,Lns2]=load4(110);
+    [Ls1,Lns1,Ls2,Lns2]=load1(110);
     historicalWorkloads = struct('Ls1', Ls1, 'Lns1', Lns1, 'Ls2', Ls2, 'Lns2', Lns2);
     priorityTables = getPriorityTables();
     constants = struct('historicalWorkloads', historicalWorkloads, 'priorityTables', priorityTables, 'generatorOutput', generatorOutput, 'nTimesteps', nTimesteps, 'Nt', Nt, 'Nl', Nl, 'Ns', Ns, 'Nb', Nb, 'N', N); %hard-coded params to pass around  
@@ -24,7 +24,7 @@ function [] = runAirplane()
     advice = [];
     HLclock = 1; %count up to each time we call the HLLMS
     for time=1:nTimesteps
-        workload = genWorkload(historicalWorkloads, time);
+        workload = genWorkload(historicalWorkloads, time, 0);
         genStatus = getGeneratorStatus(time);
         sensors = struct('workload', workload, 'genStatus', genStatus, 'time', time);
 

### 2012-12-06 project/HLLMS.m
fixed off-by-one in the outputs from Mehdi's code

@@ -74,12 +74,12 @@ function [configs] = HLLMS(sensors, constants) %only using 'sensors' for generat
     GeneratorOnOff = double(alpha(:,:)); % unlike the other variables, alpha's time index comes first
 
     configs = []; %array of 'config' data structures
-    for i=1:Nt %1 to horizon
+    for i=1:(Nt-1) %1 to horizon
         BusGen = [0 0];
-        [myMax BusGen(1)] = max(Del1_double(:,i));  %BusGen(1) is argmax here
-        [myMax BusGen(2)] = max(Del2_double(:,i));
+        [myMax BusGen(1)] = max(Del1_double(:,i+1));  %BusGen(1) is argmax here
+        [myMax BusGen(2)] = max(Del2_double(:,i+1));
 
-        config = struct('Shedding1', Shedding1(:,i)', 'Shedding2', Shedding2(:,i)', 'BusGen', BusGen, 'Battery1', Battery1(:,i), 'Battery2', Battery2(:,i), 'GeneratorOnOff', GeneratorOnOff(i,:));
+        config = struct('Shedding1', Shedding1(:,i+1)', 'Shedding2', Shedding2(:,i+1)', 'BusGen', BusGen, 'Battery1', Battery1(:,i), 'Battery2', Battery2(:,i), 'GeneratorOnOff', GeneratorOnOff(i,:));
         configs = [configs config];
     end
 end

### 2012-12-06 project/checkSafety.m

@@ -3,8 +3,8 @@
 function isSafe = checkSafety(config, sensors, constants)
     Shedding1 = config.Shedding1;
     Shedding2 = config.Shedding2;
-    Bus1_pwrReq = sum(sensors.workload.Ls1) + sum(sensors.workload.Lns1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate'
-    Bus2_pwrReq = sum(sensors.workload.Ls2) + sum(sensors.workload.Lns2(Shedding2>0));
+    Bus1_pwrReq = sum(sensors.workload.Lns1) + sum(sensors.workload.Ls1(Shedding1>0)); %'Shedding1>0' is a Matlab 'logical coordinate'
+    Bus2_pwrReq = sum(sensors.workload.Lns2) + sum(sensors.workload.Ls2(Shedding2>0));
    
     %calculate pwr required from each generator
     pwrReqGen1 = 0; pwrReqGen2 = 0; pwrReqApu = 0; 

### 2012-12-06 project/runAirplane.m

@@ -16,7 +16,7 @@ function [] = runAirplane()
     generatorOutput = [1e5, 1e5, 104e3]; %Pwr produced by generators. Called U1, U2, U3 in Mehdi's code
         %TODO: make generatorOutput be a parameter to Mehdi's code, so that we can tweak it easily.
     %[Ls1,Lns1,Ls2,Lns2]=load3(typecast(nTimesteps+Nt, 'int32')); % load the "loads" -- choose between load1, load2 and load3. (including extra padding, Nt, to avoid out-of-bounds in HLLMS)
-    [Ls1,Lns1,Ls2,Lns2]=load1(110);
+    [Ls1,Lns1,Ls2,Lns2]=load3(110);
     historicalWorkloads = struct('Ls1', Ls1, 'Lns1', Lns1, 'Ls2', Ls2, 'Lns2', Lns2);
     priorityTables = getPriorityTables();
     constants = struct('historicalWorkloads', historicalWorkloads, 'priorityTables', priorityTables, 'generatorOutput', generatorOutput, 'nTimesteps', nTimesteps, 'Nt', Nt, 'Nl', Nl, 'Ns', Ns, 'Nb', Nb, 'N', N); %hard-coded params to pass around  

### 2012-12-06 project/report/paper.tex
modifying plots so that all the lines don't overlap
refining the paper

@@
 
-\title{OL-LMS}
+\title{Hierarchical Control for Aircraft Electronic Power Systems}
 
@@
 
-\keywords{embedded systems and planes and stuff}
+\keywords{avionics, control, optimization}
 
 \section{Introduction}
-Overview, stuff about platform based design...
+High-level... Increasing complexity of EPS...
 
+We aim to design a Load Management System (LMS) for EPS that is ``smarter" in terms of several criteria.
+First, always preserve safety.
+After preserving safety, maximize the quality of service.
+Finally, conserve fuel by using as few generators as possible.
+
+The rest of the paper is organized as follows. (TODO)
+
+\section{Aircraft EPS Overview}
 Intro to aircraft EPS... explain the buses and generators.
@@
 
-Our key objective is to design a Load Management System (LMS) for EPS that is ``smarter" in terms of several criteria.
-First, always preserve safety.
-After preserving safety, maximize the quality of service.
-Finally, conserve fuel by using as few generators as possible.
-
+\subsection{Load Shedding}
 Aircraft EPS generally have two types of loads: sheddable and non-sheddable.
@@
 If an EPS power bus requires more power than the generators can provide, the LMS is responsible for disabling (shedding) some of the sheddable loads.
-
-\subsection{OL-LMS Goals}
-Talk about goals of project
 
@@
 
-\subsection{Optimizing-LMS}
+\section{Optimizing LMS}
 \label{sec:optimizing-LMS}
@@
 The key insight in [Mehdi's paper] is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
-Then, [Mehdi's paper] selected an objective function such that solving this optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
+Further, [Mehdi's paper] proposes an objective function such that solving the optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
 [This optimizing controller aims to minimize three weighted parameters: load shedding preference, generator selection preference, and number of generators used.]
@@
 
-\section{OL-LMS}
-Talk about how the entire system works at a whole, how we set it up, etc.
+TODO: explain interaction and communication between Optimizing LMS and LL-LMS.
+
+\section{Experimental Design}
+
+Parameters: \\
+generator 1: 100,000 W \\
+generator 2: 100,000 W \\
+APU: 104,000 W \\
+
+[TODO: show priority tables]
+
 \section{Experimental Results}
@@
 generators failures in the system using the power usage profile shown in
-Figure~\ref{fig:preqnofail}. Notice that Bus 1 uses more than 100000 Watts
+Figure~\ref{fig:preqnofail}. Notice that Bus 1 requests more than 100,000 Watts
 between time 30 and 60 and again around time 90. Bus 2, on the other hand,

### 2012-12-07 project/report/paper.tex
rewording a couple things in results section

@@
 For our first experiment, we simulated the case when there are no power
-generators failures in the system using the power usage profile shown in
-Figure~\ref{fig:preqnofail}. Notice that Bus 1 requests more than 100,000 Watts
+generator failures in the system using the power usage profile shown in
+Figure~\ref{fig:preqnofail}. 
+In other words, the real power usage profile (Figure~\ref{fig:preqnofail}) is given to the Optimizing LMS and LL-LMS without any perturbations.
+Notice that Bus 1 requests more than 100,000 Watts
 between time 30 and 60 and again around time 90. Bus 2, on the other hand,
 has a massive power usage between time 50 and 80. We wanted to see how the
-HL-LMS and the LL-LMS responds to this scenario in isolation. Our results
-are in Figure~\ref{fig:lshlnofail} and Figure~\ref{fig:lsllnofail}.
+HL-LMS and the LL-LMS responds to this scenario in isolation. 
+%Our results are in Figure~\ref{fig:lshlnofail} and Figure~\ref{fig:lsllnofail}.
 Comparing the results, we can see that the LL-LMS will shed a much larger
-number of loads than the the HL-LMS. In particular, the LL-LMS sheds
-exclusively from load 6-10 because it simply examines the priority table.
-The HL-LMS; however, is much smarter. It is able to minimize the number of
+number of loads (Figure~\ref{fig:lshlnofail}) than the the HL-LMS (Figure~\ref{fig:lsllnofail}). 
+In particular, the LL-LMS sheds exclusively from load 6-10 because it simply examines the priority table.
+However, the HL-LMS however is much smarter. It is able to minimize the number of
 loads shed by shedding loads much higher up in the priority table.
@@
 \subsection{Generator Failure}
-For this experiment, we wanted to see how a combined HL-LMS and LL-LLMS
-system would respond to spontaneous generator failures. To this end, we
+Now, we investigate how our Hierarchical LMS system responds to spontaneous generator failures. 
+To this end, we
 simulated the case when generator 2 would fail at time 45. Obviously, if the HL-LMS were
@@
 LL-LMS system. The HL-LMS produces advice to be used from time 40 to 50. Unfortunately,
-there is no way for this advice to be aware of the failure at time 45. So while the advice
+there is no way for the Optimizing LMS's advice to be aware of the failure at time 45. So, while the advice
 from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. The LL-LMS is

### 2012-12-07 project/report/paper.tex
HL-LMS -> Optimizing LMS

@@
   \includegraphics[width=0.9\columnwidth]{figures/lshlnofail.png}
-  \caption{\textbf{HL-LMS Load Shedding} Notice the long
+  \caption{\textbf{Optimizing LMS Load Shedding} Notice the long
   period of load shedding around the power usage spikes.}
@@
   \caption{\textbf{LL-LMS Load Shedding} Notice that the 
-  LL-LMS sheds much more load than the HL-LMS.}
+  LL-LMS sheds much more load than the Optimizing LMS.}
   \label{fig:lsllnofail}
@@
 has a massive power usage between time 50 and 80. We wanted to see how the
-HL-LMS and the LL-LMS responds to this scenario in isolation. 
+Optimizing LMS and the LL-LMS responds to this scenario in isolation. 
 %Our results are in Figure~\ref{fig:lshlnofail} and Figure~\ref{fig:lsllnofail}.
 Comparing the results, we can see that the LL-LMS will shed a much larger
-number of loads (Figure~\ref{fig:lshlnofail}) than the the HL-LMS (Figure~\ref{fig:lsllnofail}). 
-In particular, the LL-LMS sheds exclusively from load 6-10 because it simply examines the priority table.
-However, the HL-LMS however is much smarter. It is able to minimize the number of
-loads shed by shedding loads much higher up in the priority table.
+number of loads (Figure~\ref{fig:lshlnofail}) than the the Optimizing LMS (Figure~\ref{fig:lsllnofail}). 
+In particular, the LL-LMS sheds exclusively from loads 6-10 due to the configuration of the priority table.
+However, the Optimizing LMS however is much smarter. 
+It is able to minimize the number of loads shed by shedding loads much higher up in the priority table.
 
@@
 To this end, we
-simulated the case when generator 2 would fail at time 45. Obviously, if the HL-LMS were
+simulated the case when generator 2 would fail at time 45. Obviously, if the Optimizing LMS were
 to run in isolation, it would leave the system unpowered from time 45 to 50. 
@@
 to stay under generator 1's power rating, the LL-LMS sheds a large number of loads on the 
-second bus. Figure~\ref{fig:lsolonefail} shows what would happen in a combined HL-LMS
-LL-LMS system. The HL-LMS produces advice to be used from time 40 to 50. Unfortunately,
+second bus. Figure~\ref{fig:lsolonefail} shows what would happen in a combined Optimizing LMS
+LL-LMS system. The Optimizing LMS produces advice to be used from time 40 to 50. Unfortunately,
 there is no way for the Optimizing LMS's advice to be aware of the failure at time 45. So, while the advice
 from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. The LL-LMS is
-able to see this at time 45 and kicks in, ignoring the unsafe advice from the HL-LMS. This
+able to see this at time 45 and kicks in, ignoring the unsafe advice from the Optimizing LMS. This
 behavior is observed in Figure~\ref{fig:lsolonefail} where from time 45 to 49, a large
-number of loads are shed. Then at time 50, the HL-LMS reoptimizes by turning on the APU.
+number of loads are shed. Then at time 50, the Optimizing LMS reoptimizes by turning on the APU.
 Notice the difference in the number of loads shed onward from time 50 in 
@@
   \caption{\textbf{OL-LMS Load Shedding -- One Failure}. Notice the
-  large number of loads dropped from time 45 to 50 before the HL-LMS can reoptimize.}
+  large number of loads dropped from time 45 to 50 before the Optimizing LMS can reoptimize.}
   \label{fig:lsolonefail}
@@
 \subsection{Unexpected Power Spikes}
-The HL-LMS performs a live optimization using historical data for load power requests from
+The Optimizing LMS performs a live optimization using historical data for load power requests from
 previous flights. Due to the nature of the optimization problem, it is impossible to
 program the optimization problem with live data. As a result, it is possible for the
-HL-LMS to provide unsafe advice if there is a power spike. More specifically, it is possible
-for the HL-LMS to generate a load shedding configuration that could have one of the power
+Optimizing LMS to provide unsafe advice if there is a power spike. More specifically, it is possible
+for the Optimizing LMS to generate a load shedding configuration that could have one of the power
 generators driving a higher power request than its rating. We want to simulate such a
@@
 Figure~\ref{fig:preqnofail}. However, the actual power request for this simulation is shown
-in Figure~\ref{fig:preqpwrspike}. Without the LL-LMS, the HL-LMS would use the load
+in Figure~\ref{fig:preqpwrspike}. Without the LL-LMS, the Optimizing LMS would use the load
 shedding configuration seen in Figure~\ref{fig:lshlnofail}. Notice that there is a power
-spike from time 10 to 40 that the historical data does not account for and thus the HL-LMS
+spike from time 10 to 40 that the historical data does not account for and thus the Optimizing LMS
 is unaware of. As a result, its load shedding configuration would overload the power
@@
   \caption{\textbf{OL-LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
-  use the advice from the HL-LMS. The only time it ignores the advice from the HL-LMS is
-  from time 10 to 40 when there is a power spike that the HL-LMS is not aware of.}
+  use the advice from the Optimizing LMS. The only time it ignores the advice from the Optimizing LMS is
+  from time 10 to 40 when there is a power spike that the Optimizing LMS is not aware of.}
   \label{fig:lsolpwrspike}

### 2012-12-07 project/report/paper.tex
OL-LMS -> Optimizing LMS

@@
   \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
-  \caption{\textbf{OL-LMS Load Shedding -- One Failure}. Notice the
+  \caption{\textbf{Hierarchical LMS Load Shedding -- One Failure}. Notice the
   large number of loads dropped from time 45 to 50 before the Optimizing LMS can reoptimize.}
@@
   \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
-  \caption{\textbf{OL-LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
+  \caption{\textbf{Hierarchical LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
   use the advice from the Optimizing LMS. The only time it ignores the advice from the Optimizing LMS is

### 2012-12-07 project/report/paper.tex
minor paper tweaks

@@
 from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. The LL-LMS is
-able to see this at time 45 and kicks in, ignoring the unsafe advice from the Optimizing LMS. This
+able to see this failure at time 45 and kicks in, ignoring the unsafe advice from the Optimizing LMS. This
 behavior is observed in Figure~\ref{fig:lsolonefail} where from time 45 to 49, a large
@@
   \caption{\textbf{LL-LMS Load Shedding -- One Failure}. Notice
-  the large amounts of load shedded in order to keep power usage below the power rating
+  the large amount of loads shedded in order to keep power usage below the power rating
   of generator 1.}

### 2012-12-07 project/report/paper.tex
added example priority table (straight from Mehdi's paper. will modify it later)

@@
 
-[TODO: show priority tables]
+[TODO: show priority tables; numerical example]
+
+\begin{table}[t]
+\caption{Load Priority Table Example} %from Mehdi's paper
+\label{T:shed} 
+\centering
+\begin{tabular}{c|cc}
+ Non-sheddable loads (W) & Sheddable loads (W) & Shed priority \\ \hline
+1000 & 1000 & 1 \\
+1000 & 5000 & 2 \\
+500 & 2000 & 3 \\
+5000 & 2000 & 4 \\
+1000 & 1000 & 5 \\
+1000 & 5000 & 6 \\
+2000 & 1000 & 7 \\
+5000 & 2000 & 8 \\
+500 & 2000 & 9 \\
+2000 & 1000 & 10 \\ \hline
+\end{tabular}
+\end{table}
 

### 2012-12-07 project/report/paper.tex
references

@@
 However, priority table solutions tend to shed more loads than necessary, to use more generators than necessary, [and be ill-defined in their tradeoffs between generator allocation and load shedding].
-Toward resolving these issues, the control systems community recently designed a model-predictive optimizing controller for aircraft EPS [Mehdi's paper].
+Toward resolving these issues, the control systems community recently designed a model-predictive optimizing controller for aircraft EPS~\cite{mehdi}.
 The key insight in [Mehdi's paper] is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
@@
 
-\section{Related Work}
-Related~\cite{Lamport:LaTeX}.
-
 \section{Conclusion}

### 2012-12-07 project/report/paper.tex
formatting

@@
 Toward resolving these issues, the control systems community recently designed a model-predictive optimizing controller for aircraft EPS~\cite{mehdi}.
-The key insight in [Mehdi's paper] is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
-Further, [Mehdi's paper] proposes an objective function such that solving the optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
+The key insight in~\cite{mehdi} is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
+Further,~\cite{mehdi} proposes an objective function such that solving the optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
 [This optimizing controller aims to minimize three weighted parameters: load shedding preference, generator selection preference, and number of generators used.]
@@
 
-However, the Optimizing LMS [Mehdi's paper] is not without shortcomings.
+However, the Optimizing LMS~\cite{mehdi} is not without shortcomings.
 Chiefly, the Optimizing LMS is computationally expensive (0.5s for 10 timesteps of optimization in CPLEX on an Intel 3930K processor).

### 2012-12-07 project/report/paper.tex
updated figures, filled out LL-LMS and H-LMS

@@
 \label{sec:optimizing-LMS}
-LMS systems that are based purely on priority tables are effective for preserving safety.
+LMS systems that are based purely on priority tables are very simplistic, allowing them to frequently sample the system. These LMS systems are thus more
+responsive to failures in the system.
 However, priority table solutions tend to shed more loads than necessary, to use more generators than necessary, [and be ill-defined in their tradeoffs between generator allocation and load shedding].
@@
 \section{Low Level LMS}
-[Talk about our LL-LMS system, its inputs/outputs, functionality, etc]
-
-This section presents a Low-Level LMS (LL-LMS) similar to the controllers discussed in Section~\ref{sec:related-work}.
-The key goal (and perhaps the only) goal of this load management system is to preserve safety if workloads change or if  generators fail.
-Therefore, using a pre-defined priority table of generator assignments and loads to shed, our LL-LMS sheds loads until the power requirement is within the generator's power budget.
-
-[TODO: 4-line pseudo-code for how we apply priority tables]
-
-Other details:
-Low-level controller samples system at every  timestep (e.g. 50ms of system time) 
-Ensure that no non-sheddable load is unpowered for more  than 50ms
-
-\section{Hierarchical LMS}
-Modify LL-LMS to call the Optimizing LMS periodically.
-LL-LMS does safety checks on results from Optimizing LMS...
-
-TODO: explain interaction and communication between Optimizing LMS and LL-LMS.
-
-\section{Experimental Design}
-
-Parameters: \\
-generator 1: 100,000 W \\
-generator 2: 100,000 W \\
-APU: 104,000 W \\
-
-[TODO: show priority tables; numerical example]
-
 \begin{table}[t]
-\caption{Load Priority Table Example} %from Mehdi's paper
+\caption{\textbf{Load Priority Table Example, NS = nonsheddable, S = sheddable} 
+Loads lower in the list have lower priority while loads higher up in the list have
+higher priority. LL-LMS will shed load starting at the lower end of the list.}
 \label{T:shed}
@@
 \begin{tabular}{c|cc}
- Non-sheddable loads (W) & Sheddable loads (W) & Shed priority \\ \hline
+ NS loads (W) & S loads (W) & Shed priority \\ \hline
 1000 & 1000 & 1 \\
@@
 
+\subsection{Overview}
+As the name suggest, the LL-LMS is implemented as the last level of controller in the system. Above all else, the LL-LMS must guarantee safe operation
+conditions \footnote{For the purposes of this paper, we only consider the following safety conditions: no load remains unpowered for more than 50ms,
+loads are not connected to unpowered generators, and no generator is connected to loads whose power request exceeds the power rating of the generator}.
+Briefly, the LL-LMS uses various sensors to frequently sample the state of the system. It then uses its sensor readings along with its
+priority tables, such as the one in Table~\ref{T:shed}, to produce a safe system configuration.
+
+\subsection{Implementation}
+\label{sec:LL-LMS}
+Every cycle, the LL-LMS reads it sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. The LL-LMS
+first filters out unhealthy generator tables and then uses its generator priority table to determine which of the remaining healthy generators it should
+use to power the system. Once it has determined what generator to use, the LL-LMS will then determine how much load it has to shed. It first sums up power
+requests of all the loads (sheddable and nonsheddable) on each bus. It then determines if the generator for each of the buses are able to power all of the
+loads. If the power request is more than a generator can provide, the LL-LMS will start shedding sheddable loads starting at the load with the lowest
+priority. This simple generator selection and load shedding scheme ensure that the LL-LMS can quickly arrive at a safe system configuration.
+
+%% This section presents a Low-Level LMS (LL-LMS) similar to the controllers discussed in Section~\ref{sec:related-work}.
+%% The key goal (and perhaps the only) goal of this load management system is to preserve safety if workloads change or if  generators fail.
+%% Therefore, using a pre-defined priority table of generator assignments and loads to shed, our LL-LMS sheds loads until the power requirement is within the generator's power budget.
+
+[TODO: 4-line pseudo-code for how we apply priority tables]
+
+Other details:
+Low-level controller samples system at every  timestep (e.g. 50ms of system time) 
+Ensure that no non-sheddable load is unpowered for more  than 50ms
+
+\section{Hierarchical LMS}
+\subsection{Overview}
+Recall that the goal of our project is not to design an LMS that is safe OR optimal. Rather, we want to build an LMS that is both safe AND optimal. To this
+end, we propose a hierarchical LMS that achieves the best of both worlds by combining an optimizing LMS with a low level LMS. The OL-LMS periodically,
+though at a slow rate, samples the status of the system and produces a system configuration that is optimal. More specifically, the OL-LMS checks the
+status of the generators and solves the optimization problem detailed in Section~\ref{sec:optimizing-LMS}. The LL-LMS collects the advice from the O-LMS
+and determines whether or not the advice is safe. If the advice is safe, the LL-LMS will actuate the system based on the O-LMS's advice. Otherwise, the
+LL-LMS will fall back to its priority table. Because failures are rare, the LL-LMS will spend most of the time actuating the system based on the O-LMS's
+advice. This hierarchical LMS (H-LMS) is optimal because the O-LMS is optimal and is safe because the LL-LMS is safe.
+
+\subsection{O-LMS Modifications}
+Instead of having the O-LMS system read the sensors directly as it would in a system that relies solely on an O-LMS, it will instead receive its system 
+status information from the LL-LMS. This communication strategy saves on communication links between sensors and the O-LMS and reuses the communication
+link between the LL-LMS and the O-LMS. We also use the O-LMS to perform live optimizations of the system. That is, instead of solving one large offline
+optimization problem use a receding horizon that covers the entire flight, we have the O-LMS periodically solve many optimization problems using a much
+smaller receding horizon. Our implementation makes the O-LMS system more resilient to generator failures. Previously, the O-LMS would be completely
+oblivious to generators that fail mid-flight. With this modification, the O-LMS will eventually see that there is a generator failure and reoptimize.
+
+\subsection{LL-LMS Modifications}
+The LL-LMS will now receive advice from the HL-LMS in addition to its other functionalities detailed in Section~\ref{sec:LL-LMS}. The LL-LMS will then
+check to see that the HL-LMS's advice is safe. Specifically, the LL-LMS will check that the HL-LMS did not recommend the use of a failed generator and that
+the HL-LMS has shed enough loads so that no generator is overloaded. In the case that the HL-LMS recommended the use of a failed generator, the LL-LMS
+will use the next best generator according to its priority table and perform the appropriate load shedding. In the case that HL-LMS advice would cause a
+generator overload, the LL-LMS will shed additional loads so that the power request on the bus is under the corresponding generator's power rating.
+
+\subsection{H-LMS Implmentation}
+
+\begin{figure*}[htb]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figures/H-LMS.pdf}
+  \caption{\textbf{H-LMS Block Diagram} The O-LMS and the LL-LMS runs at different rates. The O-LMS no longer communicates directly with the EPS sensors
+and actuators. The LL-LMS is the last level of controller and communicates directly with the EPS sensors and actuators. The LL-LMS periodically sends
+generator status to the O-LMS which the O-LMS uses in its next iteration. The O-LMS produces advice that it sends to the LL-LMS. Whether or not the O-LMS's
+advice is ever actuated on is left to the discretion of the LL-LMS.}
+  \label{fig:H-LMS}
+\end{figure*}
+
+Figure~\ref{fig:H-LMS} shows the block diagram of our hierarchical controller. The different clocks fed into the two controllers indicate that the two
+controllers operate at different frequencies. More specifically, the LL-LMS operates at a high frequency, fast enough to ensure that the system is not
+in an unsafe configuration for more than 50 ms. On the other hand, the HL-LMS runs at a much slower rate due to the complexity of the optimization problem.
+It produces a set of advice (generator usage recommendations as well as load shedding configurations) that can the LL-LMS can use until the next O-LMS 
+cycle. Notice that there is an arrow going from the the LL-LMS to th H-LMS because the LL-LMS is responsible for providing the H-LMS with information of
+the status of the generators. The actual generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the H-LMS
+if it is safe or the configuration determined by the LL-LMS's priority table if the O-LMS's advice is unsafe.
+
+\section{Experimental Design}
+
+Parameters: \\
+generator 1: 100,000 W \\
+generator 2: 100,000 W \\
+APU: 104,000 W \\
+
+[TODO: show priority tables; numerical example]
+
 \section{Experimental Results}
-\begin{figure}[hb]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/preqnofail.png}
+  \includegraphics[width=\columnwidth]{figures/preqnofail}
   \caption{\textbf{Power Request}
@@
 \subsection{No Failures}
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lshlnofail.png}
+  \includegraphics[width=\columnwidth]{figures/lshlnofail}
   \caption{\textbf{Optimizing LMS Load Shedding} Notice the long
@@
 \end{figure}
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsllnofail.png}
+  \includegraphics[width=0.9\columnwidth]{figures/lsllnofail}
   \caption{\textbf{LL-LMS Load Shedding} Notice that the 
@@
 
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsllonefail.png}
+  \includegraphics[width=0.9\columnwidth]{figures/lsllonefail}
   \caption{\textbf{LL-LMS Load Shedding -- One Failure}. Notice
@@
 \end{figure}
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
+  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail}
   \caption{\textbf{Hierarchical LMS Load Shedding -- One Failure}. Notice the
@@
 
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/preqpwrspike.png}
+  \includegraphics[width=0.9\columnwidth]{figures/preqpwrspike}
   \caption{\textbf{Actual Power Request} Notice the difference between this power
@@
 \end{figure}
-\begin{figure}[ht]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail.png}
+  \includegraphics[width=0.9\columnwidth]{figures/lsolpwrspike}
   \caption{\textbf{Hierarchical LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
-  use the advice from the Optimizing LMS. The only time it ignores the advice from the Optimizing LMS is
+  use the advice from the Optimizing LMS. The only time it ignores the advice from the Optimizaing LMS is
   from time 10 to 40 when there is a power spike that the Optimizing LMS is not aware of.}

### 2012-12-07 project/report/paper.tex
updated picture and wrote intro

@@
 \section{Introduction}
-High-level... Increasing complexity of EPS...
+Aircraft Electrical Power Systems (EPS) are systems that provide power to critical loads (landing gear, navigation systems, etc.)
+as well as non critical loads (cabin lights, coffee machine, etc). Above all else, an aircraft EPS controller must ensure that
+power is constantly provided to all critical parts of the system. Since EPS are involving into increasingly complex systems, EPS
+controllers must become ``smarter''. We tackle the challenge of designing such a controller in this project. More specifically,
+we contribute a Low Level Load Management System that closely monitors system status and ensures safety. We combine this
+controller with an Optimzing Load Management System~\cite{mehdi} to obtain a Hierarchical Load Management System that is
+both safe and optimal.
 
-We aim to design a Load Management System (LMS) for EPS that is ``smarter" in terms of several criteria.
-First, always preserve safety.
-After preserving safety, maximize the quality of service.
-Finally, conserve fuel by using as few generators as possible.
-
-The rest of the paper is organized as follows. (TODO)
+We will first give a brief overview of Aircraft EPS in order to provide the necesarry background information for understanding
+the purposes of our controller. This is followed by a detailed discussion of the Optimizaing Load Management System and the
+Low Level Load Management System. After presenting both controllers, we then show how to combine the two into a Hierarchical
+Load Management System. Finally, we show that our hierarchical controller acheives safety where the Optimizing Load Management
+System would have been unable to and acheives optimality where the Low Level Load Management System cannot.
 
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[hp]
   \centering
@@
 \subsection{No Failures}
-\begin{figure}[htb]
+\begin{figure*}[hp]
   \centering
-  \includegraphics[width=\columnwidth]{figures/lshlnofail}
+  \includegraphics[width=\textwidth]{figures/lshlnofail}
   \caption{\textbf{Optimizing LMS Load Shedding} Notice the long
@@
   \label{fig:lshlnofail}
-\end{figure}
-\begin{figure}[htb]
+\end{figure*}
+\begin{figure*}[hp]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsllnofail}
+  \includegraphics[width=0.9\textwidth]{figures/lsllnofail}
   \caption{\textbf{LL-LMS Load Shedding} Notice that the 
@@
   \label{fig:lsllnofail}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[htb]
+\begin{figure*}[hp]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsllonefail}
+  \includegraphics[width=0.9\textwidth]{figures/lsllonefail}
   \caption{\textbf{LL-LMS Load Shedding -- One Failure}. Notice
@@
   \label{fig:lsllonefail}
-\end{figure}
-\begin{figure}[htb]
+\end{figure*}
+\begin{figure*}[hp]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsolonefail}
+  \includegraphics[width=0.9\textwidth]{figures/lsolonefail}
   \caption{\textbf{Hierarchical LMS Load Shedding -- One Failure}. Notice the
@@
   \label{fig:lsolonefail}
-\end{figure}
+\end{figure*}
 
@@
 \end{figure}
-\begin{figure}[htb]
+\begin{figure*}[hp]
   \centering
-  \includegraphics[width=0.9\columnwidth]{figures/lsolpwrspike}
+  \includegraphics[width=0.9\textwidth]{figures/lsolpwrspike}
   \caption{\textbf{Hierarchical LMS Load Shedding -- Power Spike} Whenever possible the LL-LMS will
@@
   \label{fig:lsolpwrspike}
-\end{figure}
+\end{figure*}
 

### 2012-12-07 project/report/paper.tex
lots of refinements to paper

@@
 \makeatother
+\newcommand*{\bfrac}[2]{\genfrac{(}{)}{0pt}{}{#1}{#2}} %fraction without line
+
 \begin{document}
@@
 
-\keywords{avionics, control, optimization}
+\keywords{avionics, model-predictive control, optimization, linear programming}
 
@@
 as well as non critical loads (cabin lights, coffee machine, etc). Above all else, an aircraft EPS controller must ensure that
-power is constantly provided to all critical parts of the system. Since EPS are involving into increasingly complex systems, EPS
-controllers must become ``smarter''. We tackle the challenge of designing such a controller in this project. More specifically,
+power is constantly provided to all critical parts of the system. Since EPS are evolving into increasingly complex systems, EPS
+controllers must become ``smarter''. We address the challenge of designing such a controller in this project. More specifically,
 we contribute a Low Level Load Management System that closely monitors system status and ensures safety. We combine this
@@
 
-We will first give a brief overview of Aircraft EPS in order to provide the necesarry background information for understanding
-the purposes of our controller. This is followed by a detailed discussion of the Optimizaing Load Management System and the
-Low Level Load Management System. After presenting both controllers, we then show how to combine the two into a Hierarchical
-Load Management System. Finally, we show that our hierarchical controller acheives safety where the Optimizing Load Management
-System would have been unable to and acheives optimality where the Low Level Load Management System cannot.
+The rest of the paper is organized as follows.
+In Sections~\ref{sec:overview} and~\ref{sec:related-work}, we give an overview of Aircraft EPS.
+%in order to provide the necessary background information for understanding the purposes of our controller.
+This is followed by a detailed discussion of an Optimizing Load Management System (Section~\ref{sec:optimizing-LMS}) and a Low Level Load Management System (Section~\ref{sec:LL-LMS}). 
+After presenting these two controllers, we then show how to combine the two into a Hierarchical Load Management System. 
+Finally, in Sections~\ref{sec:numerical-examples}-\ref{sec:results}, we show that our hierarchical controller achieves safety where the Optimizing Load Management System would have been unable to and achieves optimality where the Low Level Load Management System cannot.
 
 \section{Aircraft EPS Overview}
+\label{sec:overview}
 Intro to aircraft EPS... explain the buses and generators.
@@
 \subsection{Load Shedding}
-Aircraft EPS generally have two types of loads: sheddable and non-sheddable.
-Examples of non-sheddable loads include navigation systems, electronic actuators, and other mission-critical devices.
+Aircraft EPS generally have two types of loads: sheddable and nonsheddable.
+Examples of nonsheddable loads include navigation systems, electronic actuators, and other mission-critical devices.
 Sheddable loads include the cabin lights, kitchen, and electrical outlets for passengers' laptop computers.
@@
 In typical aircraft EPS, the load management systems aim to ensure safety during unexpected system behavior [citation needed].
-Specifically, [these systems typically require] unsheddable loads to remain powered at all times. [or, maybe unsheddable loads can be unpowered for up to 50ms while switching generators]
-Therefore, if a generator fails, the EPS must shed loads and/or reallocate generators to buses such that no unsheddable load becomes unpowered.
+Specifically, these systems require nonsheddable loads to remain powered at all times.\footnote{More precisely, nonsheddable loads must not be more unpowered for more than one cycle of the LMS.
+In practice, 50ms is a typical LMS sampling frequency.} %[or, maybe nonsheddable loads can be unpowered for up to 50ms while switching generators]
+Therefore, if a generator fails, the EPS must shed loads and/or reallocate generators to buses such that no nonsheddable load becomes unpowered.
 Typically, these systems rely on hard-coded priority tables for selecting loads to shed and for allocating generators to buses.
@@
 \label{sec:optimizing-LMS}
-LMS systems that are based purely on priority tables are very simplistic, allowing them to frequently sample the system. These LMS systems are thus more
-responsive to failures in the system.
-However, priority table solutions tend to shed more loads than necessary, to use more generators than necessary, [and be ill-defined in their tradeoffs between generator allocation and load shedding].
-Toward resolving these issues, the control systems community recently designed a model-predictive optimizing controller for aircraft EPS~\cite{mehdi}.
-The key insight in~\cite{mehdi} is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an optimization problem.
-Further,~\cite{mehdi} proposes an objective function such that solving the optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
-[This optimizing controller aims to minimize three weighted parameters: load shedding preference, generator selection preference, and number of generators used.]
-[TODO: show a simplified version of Mehdi's math]
+LMS systems that are based purely on priority tables are very simplistic, allowing them to frequently sample the system. 
+These LMS systems are thus more responsive to failures in the system.
+However, pure priority table solutions tend to shed more loads than necessary, to use more generators than necessary, 
+%[and be ill-defined in their tradeoffs between generator allocation and load shedding].
+Toward resolving these issues, the control systems community recently designed a model-predictive Optimizing LMS for aircraft EPS~\cite{mehdi}.
+The key insight in~\cite{mehdi} is to formulate the system requirements (e.g. the maximum power produced by each generator) as constraints in an linear optimization problem.
+%Further,~\cite{mehdi} proposes an objective function such that solving the optimization problem produces a power system allocation using a minimal number of generators while also minimizing the total number and duration of loads that are shed.
+The Optimizing LMS minimizes an objective function comprised of three weighted vectors: load shedding preference, generator selection preference, and number of generators used.
+The following expression gives a high-level view of the objective function in~\cite{mehdi}, where the weights can correspond with priorities in priority tables and $h$ is a time horizon:
+
+$\int_t^{t+h} min(\bfrac{\text{sheddable}}{\text{loads}} * \text{\scriptsize \emph{weights}$_1$} +  \bfrac{\text{generator}}{\text{selection}} * \text{\scriptsize \emph{weights}$_2$} +  \bfrac{\text{\scriptsize \# of generators}}{\text{\scriptsize  used}})$
+
+Notice the integral from time $t$ to time $t+horizon$. 
+If the horizon=1, then we simply optimize the objective function for one system timestep.
+However, if the horizon=10, then we do one call to the CPLEX MILP solver to compute the optimal system configurations for 10 timesteps.
+Batching several timesteps into one MILP call can reduce the overall computation time: using a 6-core Intel 3930K processor, optimizing 1 timestep takes an average of 350ms, while optimizing 10 timesteps takes an average of 500ms (so, 50ms per timestep in the batched configuration).
+
+However, our {\emph receeding horizon} approach of doing a batched optimization at every 10 timesteps is not foolproof.
+This is because, at time $t$, we can't perfectly predict the power loads or generator failures at timesteps in the range [$t+1$, $t+horizon$]. 
+Therefore, instead of using live data from the load sensors, the Optimizing LMS uses historical data, averaged from previous flights that flew the same route using the same model of aircraft.
 
@@
 \section{Low Level LMS}
+\label{sec:LL-LMS}
 \begin{table}[t]
@@
 \subsection{Implementation}
-\label{sec:LL-LMS}
 Every cycle, the LL-LMS reads it sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. The LL-LMS
@@
 Low-level controller samples system at every  timestep (e.g. 50ms of system time) 
-Ensure that no non-sheddable load is unpowered for more  than 50ms
+Ensure that no nonsheddable load is unpowered for more  than 50ms
 
@@
 
+\section{Numerical Examples}
+\label{sec:numerical-examples}
+
 \section{Experimental Design}
@@
 \section{Experimental Results}
+\label{sec:results}
 \begin{figure}[htb]

### 2012-12-07 project/report/paper.tex
abstract

@@
 \begin{abstract}
-Best abstract ever
+As avionics and other aircraft electronics system increase in complexity, the task of controlling aircraft electronic power systems (EPS) becomes more challenging.
+With this in mind, we present an aircraft EPS controller, or \emph{load management system (LMS} that guarantees safety while aiming to optimize quality of service metrics such as minimizing load shedding.
+Our controller uses a hierarchical control to combine an optimization-based model predictive LMS with a priority table based low-level LMS.
+The fundamental contribution is in coupling the two controllers to work in a hierarchical fashion, where the optimization-based LMS looks forward and the low-level LMS guarantees safety in the short term.
+Through simulations and numerical examples, we find that our hierarchical controller minimizes load shedding while always preserving safety in the aircraft EPS.
 \end{abstract}
@@
 Therefore, instead of using live data from the load sensors, the Optimizing LMS uses historical data, averaged from previous flights that flew the same route using the same model of aircraft.
-
-However, the Optimizing LMS~\cite{mehdi} is not without shortcomings.
-Chiefly, the Optimizing LMS is computationally expensive (0.5s for 10 timesteps of optimization in CPLEX on an Intel 3930K processor).
-[Discuss receding horizon and the need for historical data]
+%However, the Optimizing LMS~\cite{mehdi} is not without shortcomings.
+%Chiefly, the Optimizing LMS is computationally expensive (0.5s for 10 timesteps of optimization in CPLEX on an Intel 3930K processor).
+%[Discuss receding horizon and the need for historical data]
 Since the Optimizing LMS relies on historical data, it may produce unsafe allocations that use a broken generator or operate above the generator power caps.

### 2012-12-07 project/report/paper.tex
quick discussion of timing

@@
 As avionics and other aircraft electronics system increase in complexity, the task of controlling aircraft electronic power systems (EPS) becomes more challenging.
-With this in mind, we present an aircraft EPS controller, or \emph{load management system (LMS} that guarantees safety while aiming to optimize quality of service metrics such as minimizing load shedding.
+With this in mind, we present an aircraft EPS controller, or \emph{load management system (LMS)}, that guarantees safety while aiming to optimize quality of service metrics such as minimizing load shedding.
 Our controller uses a hierarchical control to combine an optimization-based model predictive LMS with a priority table based low-level LMS.
@@
 
-\section{Low Level LMS}
+\section{Low Level LMS (LL-LMS)}
 \label{sec:LL-LMS}
@@
 
-\subsection{Overview}
-As the name suggest, the LL-LMS is implemented as the last level of controller in the system. Above all else, the LL-LMS must guarantee safe operation
-conditions \footnote{For the purposes of this paper, we only consider the following safety conditions: no load remains unpowered for more than 50ms,
-loads are not connected to unpowered generators, and no generator is connected to loads whose power request exceeds the power rating of the generator}.
+\subsection{LL-LMS Overview}
+As the name suggests, the LL-LMS serves as a simple, fail-safe controller in an aircraft power system.
+Above all else, the LL-LMS must guarantee safe operation conditions.\footnote{For the purposes of this paper, we consider the following safety requirements: no load remains unpowered for more than 50ms, loads are not connected to unpowered generators, and no generator is connected to loads whose power request exceeds the power rating of the generator.}
 Briefly, the LL-LMS uses various sensors to frequently sample the state of the system. It then uses its sensor readings along with its
@@
 \subsection{Implementation}
-Every cycle, the LL-LMS reads it sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. The LL-LMS
-first filters out unhealthy generator tables and then uses its generator priority table to determine which of the remaining healthy generators it should
-use to power the system. Once it has determined what generator to use, the LL-LMS will then determine how much load it has to shed. It first sums up power
-requests of all the loads (sheddable and nonsheddable) on each bus. It then determines if the generator for each of the buses are able to power all of the
-loads. If the power request is more than a generator can provide, the LL-LMS will start shedding sheddable loads starting at the load with the lowest
-priority. This simple generator selection and load shedding scheme ensure that the LL-LMS can quickly arrive at a safe system configuration.
+Every cycle, the LL-LMS reads it sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. 
+The LL-LMS first filters out unhealthy generator tables and then uses its generator priority table to determine which of the remaining healthy generators it should use to power the system. 
+Once it has determined what generator to use, the LL-LMS will then determine how much load it has to shed. 
+It first sums up power requests of all the loads (sheddable and nonsheddable) on each bus. 
+It then determines if the generator for each of the buses are able to power all of the loads. 
+If the power request is more than a generator can provide, the LL-LMS will start shedding sheddable loads starting at the load with the lowest priority. 
+This simple generator selection and load shedding scheme ensure that the LL-LMS can quickly arrive at a safe system configuration.
 
@@
 [TODO: 4-line pseudo-code for how we apply priority tables]
-
-Other details:
-Low-level controller samples system at every  timestep (e.g. 50ms of system time) 
-Ensure that no nonsheddable load is unpowered for more  than 50ms
 
@@
 APU: 104,000 W \\
+
+Recall from Section~\ref{sec:optimizing-LMS} that the Optimizing LMS can optimize over a horizon of 10 timesteps every 500ms. 
+Further, Section~\ref{sec:related-work} mentioned that 50ms is a typical sampling frequency for a LMS system.
+We design our simulation with these parameters in mind.
+Specifically, we set the LL-LMS to sample at a frequency of 50ms (controlled by the fast clock in Figure~\ref{fig:H-LMS}), we set the Optimizing-LMS to use a horizon of 10 timesteps, and thus the Optimizing-LMS gives a batch of advice to the LL-LMS every 500ms.
 

### 2012-12-08 project/report/paper.tex
more paper revisions. next: put the numerical examples (from my email) into the paper

@@
 As the name suggests, the LL-LMS serves as a simple, fail-safe controller in an aircraft power system.
-Above all else, the LL-LMS must guarantee safe operation conditions.\footnote{For the purposes of this paper, we consider the following safety requirements: no load remains unpowered for more than 50ms, loads are not connected to unpowered generators, and no generator is connected to loads whose power request exceeds the power rating of the generator.}
+Above all else, the LL-LMS must guarantee safe aircraft EPS operation.\footnote{For the purposes of this paper, we consider the following safety requirements: no load remains unpowered for more than 50ms, loads are not connected to unpowered generators, and no generator is connected to loads whose power request exceeds the power rating of the generator.}
 Briefly, the LL-LMS uses various sensors to frequently sample the state of the system. It then uses its sensor readings along with its
@@
 
-\subsection{Implementation}
-Every cycle, the LL-LMS reads it sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. 
+\subsection{LL-LMS Implementation}
+Every cycle, the LL-LMS reads sensors to collect data on generator status and power requests on all the nonsheddable and sheddable loads. 
 The LL-LMS first filters out unhealthy generator tables and then uses its generator priority table to determine which of the remaining healthy generators it should use to power the system. 
@@
 \section{Hierarchical LMS}
+In this section, we begin with an overview of our Hierarchical LMS.
+Then, we explain how we modify the Optimizing LMS (Section~\ref{sec:optimizing-LMS}) and LL-LMS (Section~\ref{sec:LL-LMS}) to accommodate our hierarchical strategy.
+Finally, at the end of this section, we describe the hierarchical strategy for communication between the LL-LMS and Optimizing LMS.
+
 \subsection{Overview}
-Recall that the goal of our project is not to design an LMS that is safe OR optimal. Rather, we want to build an LMS that is both safe AND optimal. To this
-end, we propose a hierarchical LMS that achieves the best of both worlds by combining an optimizing LMS with a low level LMS. The OL-LMS periodically,
-though at a slow rate, samples the status of the system and produces a system configuration that is optimal. More specifically, the OL-LMS checks the
-status of the generators and solves the optimization problem detailed in Section~\ref{sec:optimizing-LMS}. The LL-LMS collects the advice from the O-LMS
-and determines whether or not the advice is safe. If the advice is safe, the LL-LMS will actuate the system based on the O-LMS's advice. Otherwise, the
-LL-LMS will fall back to its priority table. Because failures are rare, the LL-LMS will spend most of the time actuating the system based on the O-LMS's
-advice. This hierarchical LMS (H-LMS) is optimal because the O-LMS is optimal and is safe because the LL-LMS is safe.
+Recall that the goal of our work is not to design an LMS that is safe OR optimal. 
+Rather, we want to build an LMS that is both safe AND optimal. 
+To this end, we propose a hierarchical LMS that achieves the best of both worlds by combining an optimizing LMS with a low level LMS.
+The Optimizing LMS periodically, though at a slow rate, receives the the generator status and produces a system configuration that is optimal in terms of historical workload data.
+%More specifically, the Optimizing LMS checks the status of the generators and solves the optimization problem detailed in Section~\ref{sec:optimizing-LMS}. 
+The LL-LMS collects the advice from the Optimizing LMS and determines whether or not the advice is safe. 
+If the advice is safe, the LL-LMS will actuate the system based on the Optimizing LMS's advice. 
+Otherwise, the LL-LMS will fall back to its priority table. 
+Because failures are rare, the LL-LMS will spend most of the time actuating the system based on the Optimizing LMS's advice. 
+This Hierarchical LMS is optimal because the Optimizing LMS is optimal and is safe because the LL-LMS is safe.
 
-\subsection{O-LMS Modifications}
-Instead of having the O-LMS system read the sensors directly as it would in a system that relies solely on an O-LMS, it will instead receive its system 
-status information from the LL-LMS. This communication strategy saves on communication links between sensors and the O-LMS and reuses the communication
-link between the LL-LMS and the O-LMS. We also use the O-LMS to perform live optimizations of the system. That is, instead of solving one large offline
-optimization problem use a receding horizon that covers the entire flight, we have the O-LMS periodically solve many optimization problems using a much
-smaller receding horizon. Our implementation makes the O-LMS system more resilient to generator failures. Previously, the O-LMS would be completely
-oblivious to generators that fail mid-flight. With this modification, the O-LMS will eventually see that there is a generator failure and reoptimize.
+\subsection{Optimizing LMS Modifications}
+Instead of having the Optimizing LMS system read the sensors directly as it would in a system that relies solely on an Optimizing LMS, it will instead receive its system status information from the LL-LMS.
+%This communication strategy saves on communication links between sensors and the Optimizing LMS and reuses the communication link between the LL-LMS and the Optimizing LMS. 
+This communication strategy allows the LL-LMS to have the final word on selecting a system configuration.
+%We also use the Optimizing LMS to perform live optimizations of the system. 
+%That is, 
+%smaller receding horizon. 
+%Most importantly, this implementation is resilient to generator failures. 
+Further, instead of solving one large offline optimization problem use a receding horizon that covers the entire flight, we have the Optimizing LMS periodically solve a batch of system configuration optimizations using small receding horizon.
+%Previously, the Optimizing LMS would be completely oblivious to generators that fail mid-flight. 
+With this modification, the Optimizing LMS will eventually see that there is a generator failure and reoptimize for the new topology.
 
@@
 
-\subsection{H-LMS Implmentation}
+\subsection{Hierarchical LMS Implmentation}
 
@@
   \includegraphics[width=0.75\textwidth]{figures/H-LMS.pdf}
-  \caption{\textbf{H-LMS Block Diagram} The O-LMS and the LL-LMS runs at different rates. The O-LMS no longer communicates directly with the EPS sensors
+  \caption{\textbf{Hierarchical LMS Block Diagram} The Optimizing LMS and the LL-LMS runs at different rates. The Optimizing LMS no longer communicates directly with the EPS sensors
 and actuators. The LL-LMS is the last level of controller and communicates directly with the EPS sensors and actuators. The LL-LMS periodically sends
-generator status to the O-LMS which the O-LMS uses in its next iteration. The O-LMS produces advice that it sends to the LL-LMS. Whether or not the O-LMS's
+generator status to the Optimizing LMS which the Optimizing LMS uses in its next iteration. The Optimizing LMS produces advice that it sends to the LL-LMS. Whether or not the Optimizing LMS's
 advice is ever actuated on is left to the discretion of the LL-LMS.}
@@
 in an unsafe configuration for more than 50 ms. On the other hand, the HL-LMS runs at a much slower rate due to the complexity of the optimization problem.
-It produces a set of advice (generator usage recommendations as well as load shedding configurations) that can the LL-LMS can use until the next O-LMS 
-cycle. Notice that there is an arrow going from the the LL-LMS to th H-LMS because the LL-LMS is responsible for providing the H-LMS with information of
-the status of the generators. The actual generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the H-LMS
-if it is safe or the configuration determined by the LL-LMS's priority table if the O-LMS's advice is unsafe.
+It produces a set of advice (generator usage recommendations as well as load shedding configurations) that can the LL-LMS can use until the next Optimizing LMS 
+cycle. Notice that there is an arrow going from the the LL-LMS to the Optimizing LMS because the LL-LMS is responsible for providing the Optimizing LMS with information of
+the status of the generators. The actual generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the Optimizing LMS
+if it is safe or the configuration determined by the LL-LMS's priority table if the Optimizing LMS's advice is unsafe.
 

### 2012-12-08 project/report/paper.tex
numerical examples

@@
 
+\subsection{Example: No failures or power spikes}
+We now present an example that illustrates how the Optimizing LMS manages to shed fewer loads than the LL-LMS. 
+In this example, there are no generator failures or power spikes (the historical data matches the live data), so the Hierarchical LMS would use the advice for the Optimizing LMS.
+
+Load Shedding Priority Table:
+Load 1: 60,000W, least preferred to shed.
+Load 2: 60,000W, second-least preferred to shed
+Load 3: 40,000W, most preferred to shed
+
+When presented with this workload, the LL-LMS purely applies the priority table, without any optimization or branch-and-bound to investigate ideal combinations of tasks to shed. 
+Thus, while simply shedding Load 2 would be sufficient to get system under its power cap, the LL-LMS walks up the priority table and sheds Loads 2 and 3.
+
+In contrast, the Optimizing-LMS views the priorities as weights instead of precedence constraints. 
+Therefore, the Optimizing-LMS sheds Load 2 and doesn't shed Load 3. 
+%This is one type of situation where the Optimizing-LMS [wins] over the LL-LMS. 
+Thus, the Optimizing LMS reduces load shedding compared to the LL-LMS.
+The Optimizing-LMS produces safe results in this example, so the Hierarchical LMS shares the Optimizing-LMS's reduction in load shedding.
+
+\subsection{Example: LL-LMS Preserves Safety in Power Spike}
+In this example, we show how the Hierarchical LMS preserves safety when the live workload data demands more power than the historical data predicts. 
+At Timestep 1 in the following tables, the live data matches the historical data, and the Optimizing LMS provides safe advice.
+However, Timestep 2 has a power spike in Load 3 that is not anticipated in the historical data. 
+Relying on the Optimizing-LMS alone at Timestep 2 would result in an unsafe configuration, where 120,000W is draw, while the generator can only produce 100,000W.
+Fortunately, the LL-LMS in our Hierarchical controller observes the Load 3 power spike in real-time, and it sheds Load 3 to maintain safety, thus maintaining a total power draw of 80,000W, which is within the generator's 100,000W power cap.
+
+Timestep 1 (live data matches historical data):
+    Load Shedding Priority Table:
+    Load 1: 40,000W, least preferred to shed.
+    Load 2: 40,000W, second-least preferred to shed
+    Load 3: 20,000W, most preferred to shed
+
+Timestep 2 (historical data):
+    Load Shedding Priority Table:
+    Load 1: 40,000W, least preferred to shed.
+    Load 2: 40,000W, second-least preferred to shed
+    Load 3: 20,000W, most preferred to shed
+
+Timestep 2 (live data):
+    Load Shedding Priority Table:
+    Load 1: 20,000W, least preferred to shed.
+    Load 2: 40,000W, second-least preferred to shed
+    Load 3: 40,000W, most preferred to shed
+
+\subsection{Example: LL-LMS handles generator failure}
+
+We use the following loads for all timesteps in this example:
+    Load Shedding Priority Table:
+    Load 1: 40,000W, least preferred to shed.
+    Load 2: 40,000W, second-least preferred to shed
+    Load 3: 20,000W, most preferred to shed
+
+Now, let Generator 1 fail at Timestep 2. 
+If we solely relied on the Optimizing-LMS, the left-side bus would continue to be powered by the (broken) Generator 1 during the time that the Optimizing-LMS reruns the optimization for the new topology.
+However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table). 
+(Recall that Generator 2 is preferred over the APU in the priority table.
+This is because, in addition to handling generator failures, the generator assignment is also valuable in situations with low power demand.
+For example, when taxiing around the airport, the aircraft can often conserve fuel by only using one generator.
+A byproduct of this policy is that, in the event of a generator failure, the priority table prefers to just use one generator instead of using the APU.)
+Some loads may need to be shed to run the whole system In typical operating conditions, both side buses remain powered. 
+Then, the Optimizing-LMS is alerted of the generator failure in its next optimization run. 
+When the MILP solver finishes solving the optimization problem, the Optimizing-LMS reports a new configuration to the LL-LMS. 
+This configuration would include assigning APU to the left bus.
+
 \section{Experimental Design}

### 2012-12-08 project/report/paper.tex
sld overview

@@
 \label{sec:overview}
-Intro to aircraft EPS... explain the buses and generators.
-Possibly show the diagram from the Honeywell patent.
-Possibly show our ``cartoon" diagram of an EPS system.
+\begin{figure*}[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/eps.png}
+  \caption{\textbf{Single Line Diagram of Aircraft Electrical Power System} Detailed model of the EPS system from the Honeywell Patent used
+  in this paper.}
+  \label{fig:sld}
+\end{figure*}
+Figure~\ref{fig:sld} shows a detailed Single Line Diagram (SLD) of an aircraft EPS. The diagram shows AC and DC generators connected to the engines
+as well as auxiliary power sources (APUs). These generators and power sources are responsible for providing power to the many critical and non-critical
+AC and DC loads as well as power conversion equipments. Buses are responsible for distributing power from the generators to the loads. Electrical switches
+called contactors route power from the generators to the buses. Current, voltage, and contactor sensors are used to monitor the status of the system. A
+typical EPS controller consists of Generator Control Units (GCUs) and Bus Power Control Unit (BPCU). The GCU is used to regulate the power output of the
+the generators as well as isolate failed generators. The BPCU is used to control the contactors thus making it responsible for routing power from the
+generators to the load.
 
@@
 \label{sec:numerical-examples}
+In the following numerical examples, we show how Hierarchical LMS harnesses the Optimizing LMS and LL-LMS to handle interesting load shedding situations and failure scenarios.
+For simplicity, let each of these examples be limited to the loads in the left side bus.
+Further, to keep the examples simple, we just look at sheddable loads powered by a generator with a power cap of 100,000W.
+In practice, aircraft EPS typically have a mix of sheddable and nonsheddable loads.
 
@@
 
-Load Shedding Priority Table:
-Load 1: 60,000W, least preferred to shed.
-Load 2: 60,000W, second-least preferred to shed
-Load 3: 40,000W, most preferred to shed
+%\begin{table}[t]
+%\caption{Optimizing LMS reduces load shedding compared to the LL-LMS load shedding policy.}
+%\label{T:example-1}
+%\centering
+\begin{tabular}{c|cc}
+Sheddable loads (W) & Shed priority \\ \hline
+60,000 & 1 (least preferred to shed) \\
+60,000 & 2 (middle priority to shed) \\ 
+40,000 & 3 (most preferred to shed) \\ \hline
+\end{tabular}
+%\end{table}
 

### 2012-12-08 project/report/paper.tex
priority table for example 2

@@
 
-%\begin{table}[t]
-%\caption{Optimizing LMS reduces load shedding compared to the LL-LMS load shedding policy.}
-%\label{T:example-1}
-%\centering
 \begin{tabular}{c|cc}
@@
 \end{tabular}
-%\end{table}
 
@@
 
-Timestep 1 (live data matches historical data):
-    Load Shedding Priority Table:
-    Load 1: 40,000W, least preferred to shed.
-    Load 2: 40,000W, second-least preferred to shed
-    Load 3: 20,000W, most preferred to shed
+Timestep 1 (live data matches historical data):\\
+\begin{tabular}{c|cc}
+Sheddable loads (W) & Shed priority \\ \hline
+40,000 & 1 (least preferred to shed) \\
+40,000 & 2 (middle priority to shed) \\ 
+{\bf 20,000} & 3 (most preferred to shed) \\ \hline
+\end{tabular}
 
-Timestep 2 (historical data):
-    Load Shedding Priority Table:
-    Load 1: 40,000W, least preferred to shed.
-    Load 2: 40,000W, second-least preferred to shed
-    Load 3: 20,000W, most preferred to shed
+Timestep 2 (historical data): \\
+\begin{tabular}{c|cc}
+Sheddable loads (W) & Shed priority \\ \hline
+40,000 & 1 (least preferred to shed) \\
+40,000 & 2 (middle priority to shed) \\ 
+{\bf 20,000} & 3 (most preferred to shed) \\ \hline
+\end{tabular}
 
-Timestep 2 (live data):
-    Load Shedding Priority Table:
-    Load 1: 20,000W, least preferred to shed.
-    Load 2: 40,000W, second-least preferred to shed
-    Load 3: 40,000W, most preferred to shed
+Timestep 2 (live data): \\
+\begin{tabular}{c|cc}
+Sheddable loads (W) & Shed priority \\ \hline
+40,000 & 1 (least preferred to shed) \\
+40,000 & 2 (middle priority to shed) \\ 
+{\bf 40,000} & 3 (most preferred to shed) \\ \hline
+\end{tabular}
 

### 2012-12-08 project/report/paper.tex
adding some details to the 'Hierarchical LMS implementation' subsection:

--- 2012-12-08 project/report/paper.tex
@@
 
-\subsection{Optimizing LMS Modifications}
+\subsection{Modifications to Optimizing LMS}
 Instead of having the Optimizing LMS system read the sensors directly as it would in a system that relies solely on an Optimizing LMS, it will instead receive its system status information from the LL-LMS.
@@
 
-\subsection{LL-LMS Modifications}
-The LL-LMS will now receive advice from the HL-LMS in addition to its other functionalities detailed in Section~\ref{sec:LL-LMS}. The LL-LMS will then
-check to see that the HL-LMS's advice is safe. Specifically, the LL-LMS will check that the HL-LMS did not recommend the use of a failed generator and that
-the HL-LMS has shed enough loads so that no generator is overloaded. In the case that the HL-LMS recommended the use of a failed generator, the LL-LMS
-will use the next best generator according to its priority table and perform the appropriate load shedding. In the case that HL-LMS advice would cause a
+\subsection{Modifications to LL-LMS}
+The LL-LMS will now receive advice from the Optimizing LMS in addition to its other functionalities detailed in Section~\ref{sec:LL-LMS}. The LL-LMS will then
+check to see that the Optimizing LMS's advice is safe. Specifically, the LL-LMS will check that the Optimizing LMS did not recommend the use of a failed generator and that
+the Optimizing LMS has shed enough loads so that no generator is overloaded. In the case that the Optimizing LMS recommended the use of a failed generator, the LL-LMS
+will use the next best generator according to its priority table and perform the appropriate load shedding. In the case that Optimizing LMS advice would cause a
 generator overload, the LL-LMS will shed additional loads so that the power request on the bus is under the corresponding generator's power rating.
@@
 
-Figure~\ref{fig:H-LMS} shows the block diagram of our hierarchical controller. The different clocks fed into the two controllers indicate that the two
-controllers operate at different frequencies. More specifically, the LL-LMS operates at a high frequency, fast enough to ensure that the system is not
-in an unsafe configuration for more than 50 ms. On the other hand, the HL-LMS runs at a much slower rate due to the complexity of the optimization problem.
-It produces a set of advice (generator usage recommendations as well as load shedding configurations) that can the LL-LMS can use until the next Optimizing LMS 
-cycle. Notice that there is an arrow going from the the LL-LMS to the Optimizing LMS because the LL-LMS is responsible for providing the Optimizing LMS with information of
-the status of the generators. The actual generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the Optimizing LMS
+Figure~\ref{fig:H-LMS} shows the block diagram of our hierarchical controller. 
+The different clocks fed into the two controllers indicate that the two controllers operate at different frequencies. 
+More specifically, the LL-LMS operates at a high frequency, fast enough to ensure that the system is not in an unsafe configuration for more than 50 ms.
+
+On the other hand, the Optimizing LMS runs at a much slower rate due to the complexity of the optimization problem.
+Recall from Section~\ref{sec:optimizing-LMS} that the Optimizing LMS can optimize over a horizon of 10 timesteps every 500ms. 
+Therefore, we set the LL-LMS to sample at a frequency of 50ms (controlled by the fast clock in Figure~\ref{fig:H-LMS}), we set the Optimizing-LMS to use a horizon of 10 timesteps, and thus the Optimizing-LMS gives a batch of advice to the LL-LMS every 500ms.
+
+At each 500ms interval, the Optimizing LMS produces a set of 10 timesteps worth of advice (generator usage recommendations as well as load shedding configurations) that can the LL-LMS can use until the next Optimizing LMS cycle.
+Notice that there is an arrow going from the the LL-LMS to the Optimizing LMS because the LL-LMS is responsible for providing the Optimizing LMS with information of the status of the generators. 
+The generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the Optimizing LMS
 if it is safe or the configuration determined by the LL-LMS's priority table if the Optimizing LMS's advice is unsafe.
@@
 
-We use the following loads for all timesteps in this example:
-    Load Shedding Priority Table:
-    Load 1: 40,000W, least preferred to shed.
-    Load 2: 40,000W, second-least preferred to shed
-    Load 3: 20,000W, most preferred to shed
+We use the following loads for all timesteps in this example: \\
+\begin{tabular}{c|cc}
+Sheddable loads (W) & Shed priority \\ \hline
+40,000 & 1 (least preferred to shed) \\
+40,000 & 2 (middle priority to shed) \\ 
+20,000 & 3 (most preferred to shed) \\ \hline
+\end{tabular}
 
+As with our other examples in this section, we're concentrating on the left side bus, which is powered by the left generator (labeled as Generator 1).
 Now, let Generator 1 fail at Timestep 2. 
 If we solely relied on the Optimizing-LMS, the left-side bus would continue to be powered by the (broken) Generator 1 during the time that the Optimizing-LMS reruns the optimization for the new topology.
-However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table). 
-(Recall that Generator 2 is preferred over the APU in the priority table.
+However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table).\footnote{Recall that Generator 2 is preferred over the APU in the left side bus's priority table.
 This is because, in addition to handling generator failures, the generator assignment is also valuable in situations with low power demand.
 For example, when taxiing around the airport, the aircraft can often conserve fuel by only using one generator.
-A byproduct of this policy is that, in the event of a generator failure, the priority table prefers to just use one generator instead of using the APU.)
-Some loads may need to be shed to run the whole system In typical operating conditions, both side buses remain powered. 
+A byproduct of this policy is that, in the event of a generator failure, the priority table prefers to run the entire aircraft on a single generator instead of using the APU.}
+Some loads may need to be shed to run the whole system off of generator 2.
+However, so long as the nonsheddable loads in both buses are within the power cap of generator 2 then, both side buses remain powered. 
 Then, the Optimizing-LMS is alerted of the generator failure in its next optimization run. 
 When the MILP solver finishes solving the optimization problem, the Optimizing-LMS reports a new configuration to the LL-LMS. 
-This configuration would include assigning APU to the left bus.
+This configuration would include assigning the APU to power the left bus.
 
@@
 APU: 104,000 W \\
-
-Recall from Section~\ref{sec:optimizing-LMS} that the Optimizing LMS can optimize over a horizon of 10 timesteps every 500ms. 
-Further, Section~\ref{sec:related-work} mentioned that 50ms is a typical sampling frequency for a LMS system.
-We design our simulation with these parameters in mind.
-Specifically, we set the LL-LMS to sample at a frequency of 50ms (controlled by the fast clock in Figure~\ref{fig:H-LMS}), we set the Optimizing-LMS to use a horizon of 10 timesteps, and thus the Optimizing-LMS gives a batch of advice to the LL-LMS every 500ms.
 

### 2012-12-08 project/report/paper.tex
extended the 'describing aircraft EPS' section

@@
 generators to the load.
+
+In this paper, we discuss \emph{load management systems (LMS)} that encompass the GCU and BPCU responsibilities such as load shedding and the assignment of generators to buses. 
+The load shedding and generator assignments are actuated by the contactors, though we focus on the high-level control and assume there is an interface to convert our control instructions to contactor configurations.
+
+Also, most of the examples in this paper use a system with the following configuration.
+Each side of the aircraft has a bus, and the left-side and right-side buses each have a set of sheddable and nonsheddable tasks.
+There are three generators--Generator 1 is part of the left-side engine, Generator 2 is part of the right-side engine, and the auxiliary power unit (APU) is generally used only when one or both of the main generators has failed.
 
@@
 
-[TODO: 4-line pseudo-code for how we apply priority tables]
+%[TODO: 4-line pseudo-code for how we apply priority tables]
 

### 2012-12-08 project/report/paper.tex
test
generator pri table
added ref to the generator pri table

@@
 2000 & 1000 & 10 \\ \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[t]
+\caption{Priority table for allocating generators to buses.}
+\label{T:generator-priority}
+\centering
+\begin{tabular}{c|cc}
+ Priority & Left bus & Right bus \\ \hline
+1 (most preferred) & Generator 1 & Generator 2 \\
+2 ~(mid preferred) & Generator 2 & Generator 1 \\
+3 (least preferred) & APU & APU \\ \hline
 \end{tabular}
@@
 If we solely relied on the Optimizing-LMS, the left-side bus would continue to be powered by the (broken) Generator 1 during the time that the Optimizing-LMS reruns the optimization for the new topology.
-However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table).\footnote{Recall that Generator 2 is preferred over the APU in the left side bus's priority table.
+However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table).\footnote{Recall from Table~\ref{T:generator-priority} that Generator 2 is preferred over the APU in the left side bus's priority table.
 This is because, in addition to handling generator failures, the generator assignment is also valuable in situations with low power demand.

### 2012-12-08 project/report/paper.tex
doing a pass through the results section

@@
 If we solely relied on the Optimizing-LMS, the left-side bus would continue to be powered by the (broken) Generator 1 during the time that the Optimizing-LMS reruns the optimization for the new topology.
-However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priority table).\footnote{Recall from Table~\ref{T:generator-priority} that Generator 2 is preferred over the APU in the left side bus's priority table.
+However, in the Hierarchical LMS, the LL-LMS steps in at the time of the generator failure and reassigns the left side bus to Generator 2 (following the priorities in Table~\ref{T:generator-priority}).\footnote{Notice in Table~\ref{T:generator-priority} that Generator 2 is preferred over the APU in the left side bus's priority table.
 This is because, in addition to handling generator failures, the generator assignment is also valuable in situations with low power demand.
@@
 This configuration would include assigning the APU to power the left bus.
-
-\section{Experimental Design}
-
-Parameters: \\
-generator 1: 100,000 W \\
-generator 2: 100,000 W \\
-APU: 104,000 W \\
-
-[TODO: show priority tables; numerical example]
 
@@
 \subsection{Unexpected Power Spikes}
-The Optimizing LMS performs a live optimization using historical data for load power requests from
-previous flights. Due to the nature of the optimization problem, it is impossible to
-program the optimization problem with live data. As a result, it is possible for the
-Optimizing LMS to provide unsafe advice if there is a power spike. More specifically, it is possible
-for the Optimizing LMS to generate a load shedding configuration that could have one of the power
-generators driving a higher power request than its rating. We want to simulate such a
-failure case and show that the LL-LMS performs the required adjustment to keep the system
-in a safe state. For this experiment, we use the same historical power request as in
-Figure~\ref{fig:preqnofail}. However, the actual power request for this simulation is shown
-in Figure~\ref{fig:preqpwrspike}. Without the LL-LMS, the Optimizing LMS would use the load
-shedding configuration seen in Figure~\ref{fig:lshlnofail}. Notice that there is a power
-spike from time 10 to 40 that the historical data does not account for and thus the Optimizing LMS
-is unaware of. As a result, its load shedding configuration would overload the power
-generator. Because we have a LL-LMS, we are able to keep the system in a safe configuration.
-Figure~\ref{fig:lsolpwrspike} shows that the LL-LMS will detect that the generators would
-be overloaded and thus shed some loads during this time interval.
+\label{sec:power-spikes-simulation}
+The Optimizing LMS performs a live optimization using historical data for load power requests from previous flights. 
+Due to the nature of the optimization problem, it is impossible to program the optimization problem with live data. 
+As a result, it is possible for the Optimizing LMS to provide unsafe advice if there is a power spike. 
+More specifically, it is possible for the Optimizing LMS to generate a load shedding configuration that could have one of the power generators driving a higher power request than its rating.
+We want to simulate such a failure case and show that the LL-LMS performs the required adjustment to keep the system in a safe state.
+For this experiment, we use the same historical power request as in Figure~\ref{fig:preqnofail}. 
+However, the actual power request for this simulation is shown in Figure~\ref{fig:preqpwrspike}. 
+Without the LL-LMS, the Optimizing LMS would use the load shedding configuration seen in Figure~\ref{fig:lshlnofail}. 
+Notice that there is a power spike from time 10 to 40 that the historical data does not account for and thus the Optimizing LMS is unaware of. 
+As a result, its load shedding configuration would overload the power generator. 
+Because we have a LL-LMS, we are able to keep the system in a safe configuration.
+Figure~\ref{fig:lsolpwrspike} shows that the LL-LMS will detect that the generators would be overloaded and thus shed some loads during this time interval.
 
@@
   \includegraphics[width=0.9\columnwidth]{figures/preqpwrspike}
-  \caption{\textbf{Actual Power Request} Notice the difference between this power
+  \caption{\textbf{Actual Power Request for Section~\ref{sec:power-spikes-simulation}}. Notice the difference between this power
   request and the power request expected by the historical data in 

### 2012-12-08 project/report/paper.tex
more results section revisions

@@
 \subsection{Example: LL-LMS Preserves Safety in Power Spike}
+\label{sec:power-spike-example}
 In this example, we show how the Hierarchical LMS preserves safety when the live workload data demands more power than the historical data predicts. 
@@
 
-For our first experiment, we simulated the case when there are no power
+We begin by simulating the case when there are no power
 generator failures in the system using the power usage profile shown in
@@
 \label{sec:power-spikes-simulation}
-The Optimizing LMS performs a live optimization using historical data for load power requests from previous flights. 
-Due to the nature of the optimization problem, it is impossible to program the optimization problem with live data. 
-As a result, it is possible for the Optimizing LMS to provide unsafe advice if there is a power spike. 
-More specifically, it is possible for the Optimizing LMS to generate a load shedding configuration that could have one of the power generators driving a higher power request than its rating.
+%The Optimizing LMS performs a live optimization using historical data for load power requests from previous flights. 
+%Due to the nature of the optimization problem, it is impossible to program the optimization problem with live data. 
+As we showed in Section~\ref{sec:power-spike-example}, it is possible for the Optimizing LMS to provide unsafe advice if there is a power spike. 
+More specifically, due to the Optimizing LMS's dependence on historical data, the Optimizing LMS can generate a load shedding configuration that draws more power than the generators can produce. 
 We want to simulate such a failure case and show that the LL-LMS performs the required adjustment to keep the system in a safe state. 
-For this experiment, we use the same historical power request as in Figure~\ref{fig:preqnofail}. 
-However, the actual power request for this simulation is shown in Figure~\ref{fig:preqpwrspike}. 
+For this experiment, the Optimizing LMS is given the historical power loads in Figure~\ref{fig:preqnofail}. 
+However, the actual power loads for this simulation is shown in Figure~\ref{fig:preqpwrspike}. 
 Without the LL-LMS, the Optimizing LMS would use the load shedding configuration seen in Figure~\ref{fig:lshlnofail}. 
@@
 Because we have a LL-LMS, we are able to keep the system in a safe configuration.
-Figure~\ref{fig:lsolpwrspike} shows that the LL-LMS will detect that the generators would be overloaded and thus shed some loads during this time interval.
+Figure~\ref{fig:lsolpwrspike} shows that the LL-LMS detects that the Optimizing LMS's configuration overloads some generators, and the LL-LMS applies the priority tables to shed some loads during this time interval.
 

### 2012-12-08 project/report/paper.tex
quick intro in the experimental results section

@@
 \section{Experimental Results}
+In this section, we simulate a number of system configurations that demonstrate the Hierarchical LMS's combination of safe operation and optimization where feasible.
+The examples in Section~\ref{sec:numerical-examples} provide the intuition for some of these simulation experiments, and the simulations help to further validate the Hierarchical LMS.
+Our simulation setup is constructed in Matlab and Simulink, and the Optimizing LMS uses CPLEX to solve its optimization problem.
+
 \label{sec:results}

### 2012-12-08 project/report/paper.tex
few more small tweaks

@@
 In this section, we begin with an overview of our Hierarchical LMS.
-Then, we explain how we modify the Optimizing LMS (Section~\ref{sec:optimizing-LMS}) and LL-LMS (Section~\ref{sec:LL-LMS}) to accommodate our hierarchical strategy.
+Then, we explain how we modify the Optimizing LMS (from Section~\ref{sec:optimizing-LMS}) and LL-LMS (from Section~\ref{sec:LL-LMS}) to accommodate our hierarchical strategy.
 Finally, at the end of this section, we describe the hierarchical strategy for communication between the LL-LMS and Optimizing LMS.
@@
 \subsection{Example: No failures or power spikes}
+\label{sec:load-shedding-example}
 We now present an example that illustrates how the Optimizing LMS manages to shed fewer loads than the LL-LMS. 
@@
 In particular, the LL-LMS sheds exclusively from loads 6-10 due to the configuration of the priority table.
-However, the Optimizing LMS however is much smarter. 
-It is able to minimize the number of loads shed by shedding loads much higher up in the priority table.
+%However, the Optimizing LMS however is much smarter. 
+However, the Optimizing LMS sheds fewer loads, due to the intuition described in the numerical example in Section~\ref{sec:load-shedding-example}.
+Specifically, the Optimizing LMS reduces the number of loads shed by shedding loads by cherry-picking weighted loads throughout the priority table instead of shedding all low-priority loads before shedding high-priority loads.
+Since the system in this example experienced no generator failures or unexpected power spikes, running the system with only the Optimizing LMS or with the Hierarchical LMS produces the same result (Figure~\ref{fig:lshlnofail}).
 
@@
 Now, we investigate how our Hierarchical LMS system responds to spontaneous generator failures. 
-To this end, we
-simulated the case when generator 2 would fail at time 45. Obviously, if the Optimizing LMS were
-to run in isolation, it would leave the system unpowered from time 45 to 50. 
-Figure~\ref{fig:lsllonefail} shows what would happen if the LL-LMS were to run in 
-isolation. Starting at time 45, the LL-LMS sees that generator 2 has failed and it has 
-switched to using generator 1 to power both buses according to its priority table. In order
-to stay under generator 1's power rating, the LL-LMS sheds a large number of loads on the 
-second bus. Figure~\ref{fig:lsolonefail} shows what would happen in a combined Optimizing LMS
-LL-LMS system. The Optimizing LMS produces advice to be used from time 40 to 50. Unfortunately,
-there is no way for the Optimizing LMS's advice to be aware of the failure at time 45. So, while the advice
-from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. The LL-LMS is
-able to see this failure at time 45 and kicks in, ignoring the unsafe advice from the Optimizing LMS. This
-behavior is observed in Figure~\ref{fig:lsolonefail} where from time 45 to 49, a large
-number of loads are shed. Then at time 50, the Optimizing LMS reoptimizes by turning on the APU.
-Notice the difference in the number of loads shed onward from time 50 in 
-Figure~\ref{fig:lsllonefail} and Figure~\ref{fig:lsolonefail}.
+To this end, we simulated the case when generator 2 (which was powering the right bus) would fail at time 45. 
+Obviously, if the Optimizing LMS were to run in isolation, it would leave the system unpowered from time 45 to 50. 
+Figure~\ref{fig:lsllonefail} shows what would happen if the LL-LMS were to run in  isolation. 
+Starting at time 45, the LL-LMS sees that generator 2 has failed and it has switched to using generator 1 to power both buses according to its priority table. 
+In order to stay under generator 1's power rating, the LL-LMS sheds a large number of loads on the second bus.
+Figure~\ref{fig:lsolonefail} shows what would happen in a combined Optimizing LMS LL-LMS system.
+The Optimizing LMS produces advice to be used from time 40 to 50. 
+Unfortunately, there is no way for the Optimizing LMS's advice to be aware of the failure at time 45. 
+So, while the advice from time 40 to 44 is optimal and safe, the advice from 45 to 49 is unsafe. 
+The LL-LMS is able to see this failure at time 45 and kicks in, ignoring the unsafe advice from the Optimizing LMS. 
+This behavior is observed in Figure~\ref{fig:lsolonefail} where from time 45 to 49, a large number of loads are shed. 
+(Given the priorities in Table~\ref{T:generator-priority}, the LL-LMS prefers to use generator 1 instead of the APU to power the right bus.)
+Then at time 50, the Optimizing LMS reoptimizes by turning on the APU.
+Notice the difference in the number of loads shed onward after time 50 in Figure~\ref{fig:lsllonefail} and Figure~\ref{fig:lsolonefail}.
 
@@
   \includegraphics[width=0.9\textwidth]{figures/lsllonefail}
-  \caption{\textbf{LL-LMS Load Shedding -- One Failure}. Notice
+  \caption{\textbf{LL-LMS Load Shedding -- One Generator Failure}. Notice
   the large amount of loads shedded in order to keep power usage below the power rating
@@
   \includegraphics[width=0.9\textwidth]{figures/lsolonefail}
-  \caption{\textbf{Hierarchical LMS Load Shedding -- One Failure}. Notice the
+  \caption{\textbf{Hierarchical LMS Load Shedding -- One Generator Failure}. Notice the
   large number of loads dropped from time 45 to 50 before the Optimizing LMS can reoptimize.}
@@
 More specifically, due to the Optimizing LMS's dependence on historical data, the Optimizing LMS can generate a load shedding configuration that draws more power than the generators can produce. 
-We want to simulate such a failure case and show that the LL-LMS performs the required adjustment to keep the system in a safe state. 
+We now simulate such a failure case and show that the LL-LMS performs the required adjustment to keep the system in a safe state. 
 For this experiment, the Optimizing LMS is given the historical power loads in Figure~\ref{fig:preqnofail}. 

### 2012-12-08 project/report/paper.tex
future work

@@
 
-\section{Conclusion}
-Conclusion.
+\section{Conclusions and Future Work}
+We presented a Hierarchical load management system (LMS) for aircraft EPS.
+Our Hierarchical LMS minimizes load shedding and optimizes generator assignment when the loads are predicted appropriately, and the Hierarchical LMS reverts to simple priority tables during unexpected EPS loads or failures.
+We described the interface and interaction between the two controllers in the Hierarchical LMS, and we showed the intuition and effectiveness of our controller through numerical examples and simulation experiments.
+
+The present paper motivates several avenues for future work.
+First, in real-world scenarios, the Optimizing LMS may provide more relevant advice if it has more access to the live data. 
+Strategies for fusing the live data with the historical data for use in the Optimizing LMS would be a valuable contribution.
+Further, finding ways to reduce the computational complexity of the Optimizing LMS would allow the LL-LMS to request advice from the Optimizing LMS more frequently.
+These future work directions would ideally reduce the likelihood that the Hierarchical LMS falls back on the LL-LMS and its pure priority tables, thus increasing the amount of time that the system runs in an optimal configuration.
 
 \section*{Acknowledgments}
-This work was supported by the US Department of Defense (DoD) through the National Defense Science \& Engineering Graduate Fellowship (NDSEG) Program.
+This work was supported in part by the US Department of Defense (DoD) through the National Defense Science \& Engineering Graduate Fellowship (NDSEG) Program.
 

### 2012-12-08 project/report/paper.tex
fixed 'spill' of section 7.2 title
ref

@@
 If the horizon=1, then we simply optimize the objective function for one system timestep.
-However, if the horizon=10, then we do one call to the CPLEX MILP solver to compute the optimal system configurations for 10 timesteps.
+However, if the horizon=10, then we do one call to the CPLEX~\cite{CPLEX} MILP solver to compute the optimal system configurations for 10 timesteps.
 Batching several timesteps into one MILP call can reduce the overall computation time: using a 6-core Intel 3930K processor, optimizing 1 timestep takes an average of 350ms, while optimizing 10 timesteps takes an average of 500ms (so, 50ms per timestep in the batched configuration).
@@
 
-\subsection{Example: LL-LMS Preserves Safety in Power Spike}
+\subsection{Example: LL-LMS Preserves Safety in \\ Power Spike}
 \label{sec:power-spike-example}

### 2012-12-08 project/report/paper.tex
really flimsy related work section

@@
 Typically, these systems rely on hard-coded priority tables for selecting loads to shed and for allocating generators to buses.
+
+Beyond discussions of simply applying priority tables, perhaps the most closely related work in the Aircraft EPS community focuses on optimizing \emph{the EPS topology itself}~\cite{Chandrasekaran:2003, VanDriel:2006, Pinto:2010}.
+In contrast to these results, we focus on optimizing the control of load shedding and generator assignment of any given EPS topology, while also preserving safety.
+An other related direction is the work by Xu et al.~\cite{xu:2012}, who proposes correct-by-construction design of load management systems.
+While the work in~\cite{xu:2012} guarantees safety, it doesn't optimize for minimal load shedding or other similar metrics.
 

### 2012-12-08 project/report/paper.tex
fixed equation that extended over into next column
few last minute tweaks

@@
 \makeatother
+\newcommand*{\metafrac}[2]{\genfrac{}{}{0pt}{}{#1}{#2}} %fraction without line
 \newcommand*{\bfrac}[2]{\genfrac{(}{)}{0pt}{}{#1}{#2}} %fraction without line
@@
 After presenting these two controllers, we then show how to combine the two into a Hierarchical Load Management System. 
-Finally, in Sections~\ref{sec:numerical-examples}-\ref{sec:results}, we show that our hierarchical controller achieves safety where the Optimizing Load Management System would have been unable to and achieves optimality where the Low Level Load Management System cannot.
+Finally, in Sections~\ref{sec:numerical-examples} and~\ref{sec:results}, we show that our hierarchical controller achieves safety where the Optimizing Load Management System would have been unable to and achieves optimality where the Low Level Load Management System cannot.
 
@@
 \label{sec:related-work}
-In typical aircraft EPS, the load management systems aim to ensure safety during unexpected system behavior [citation needed].
+In typical aircraft EPS, the load management systems aim to ensure safety during unexpected system behavior.
 Specifically, these systems require nonsheddable loads to remain powered at all times.\footnote{More precisely, nonsheddable loads must not be more unpowered for more than one cycle of the LMS.
@@
 
-$\int_t^{t+h} min(\bfrac{\text{sheddable}}{\text{loads}} * \text{\scriptsize \emph{weights}$_1$} +  \bfrac{\text{generator}}{\text{selection}} * \text{\scriptsize \emph{weights}$_2$} +  \bfrac{\text{\scriptsize \# of generators}}{\text{\scriptsize  used}})$
+$\int_t^{t+h} min(\bfrac{\text{sheddable}}{\text{loads}} * \text{\scriptsize \emph{weights}$_1$} +  \bfrac{\text{generator}}{\text{selection}} * \text{\scriptsize \emph{weights}$_2$} +  \bfrac{\metafrac{\text{\scriptsize \# of }}{\text{\scriptsize generators}}}{\text{\scriptsize used}})$
 
@@
 
-\subsection{Example: LL-LMS Preserves Safety in \\ Power Spike}
+\subsection{Example: LL-LMS Preserves Safety \\ during Power Spike}
 \label{sec:power-spike-example}
@@
 
+~\\ %skip a line
 Timestep 1 (live data matches historical data):\\

### 2013-02-18 project/refactored_mehdi_code/OptProb_Linear_Function_cumsum_2-13-13.m
forking off old version with cumsum

@@ -1,0 +1,288 @@
+
+%return all decision variables
+function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear_Function()
+    % Choose between different load files, load1, load2, load3,...
+
+    close all hidden %get rid of old figures
+    clear all
+    %% constants
+    Nt=10+1;   % number of time steps   % can select: 10+1, 20+1, 50+1 and 100+1.
+    Nl=10;   % number of loads connected to each bus
+    Ns=3;    % number of power sources
+    Nb=2;    % number of HVAC buses
+    N=100;   % length of prediction horizon
+
+    %% load the "loads"
+    [Ls1,Lns1,Ls2,Lns2]=load3(N);   % choose between load1, load2 and load3.
+
+    %Load requirements for bus 1
+    Ls1sum=sum(Ls1,1); 
+    Lns1sum=sum(Lns1,1);
+
+    % Load requirements for bus 2
+    Ls2sum=sum(Ls2,1); 
+    Lns2sum=sum(Lns2,1); 
+
+    plotPowerReq(Ls1, Lns1, Ls2, Lns2, N)
+
+    %% Max. Power supply by Engines and APU
+    U1=1e5;     Peng1=U1*ones(1,Nt);
+    U2=1e5;     Peng2=U2*ones(1,Nt);
+    U3=104e3;   Papu=U3*ones(1,Nt);
+    U=[U1*ones(1,Nt); U2*ones(1,Nt); U3*ones(1,Nt)];
+    P=[Peng1' Peng2' Papu'];
+
+    %% Optimization problem set-up
+    tic;
+    %Coefficients (one set for Bus 1, one set for Bus 2):
+    Lambda1=[0 1 2];                    Lambda2=[1 0 2]; %generator priority table (GR, GL, APU)
+    Gamma1=1000*ones(1,Nl);             Gamma2=500*ones(1,Nl); %load shedding priority table (one value for each load at each timestep)
+    M=10; %'mu' -- weight for alpha (see eq.9 in OLMS paper)
+
+    % Decision variables (one set for Bus 1, one set for Bus 2)
[…]

### 2013-02-18 project/refactored_mehdi_code/OptProb_Linear_Function.m
using BETA as running total of battery charge, and Beta as per-timestep charge/discharge. the battery is still leaving a lot of juice on the table.

@@ -50,17 +50,17 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     %new decision variables for battery overflow
     Overflow1=sdpvar(1,Nt,'full'); Overflow2 = sdpvar(1,Nt,'full');
     wastePower1=sdpvar(1,Nt,'full'); wastePower2=sdpvar(1,Nt,'full'); %Power not used by loads
+    BETA1 = sdpvar(1,Nt,'full'); BETA2 = sdpvar(1,Nt,'full'); %cumulative battery charge. (lowercase Beta is per-timestep change in charge)
 
     % Constraints
     cons=[];
-    chargeRate = 1000; %TODO -- charge/discharge rate of 1000W per timestep. (arbitrary. will revise this once we look more carefully at specifications.)
+    %chargeRate = 1000; %TODO -- charge/discharge rate of 1000W per timestep. (arbitrary. will revise this once we look more carefully at specifications.)
     %cons=[cons, -chargeRate <= Beta1 <= chargeRate; -chargeRate <= Beta2 <= chargeRate]; %will add this later
 
     %doing a running total of battery charge
     timestep = 1; % 1sec for now. using this to convert W to Wh for battery capacity
     batteryCapacity = 70000; %Wh
-    cons = [cons, 0 <= cumsum(Beta1*timestep) <= batteryCapacity, 0 <= cumsum(Beta2*timestep) <= batteryCapacity];
-    %TODO: add starting charge level to cumsum(Beta). For now, we assume that battery charge is 0 at timestep 0.
+    cons=[cons, 0 <= BETA1 <= batteryCapacity, 0 <= BETA2 <= batteryCapacity];
     cons=[cons, wastePower1 >= 0, wastePower2 >= 0]; %loads can't siphon imaginary power by making wastePower negative
 
     for i=1:Nl-1
@@ -78,15 +78,10 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     cons=[cons,  0 <= Y1 <= Pito1', 0 <= Y2 <= Pito2'];
     cons=[cons, Pito1' - U.*(1-Del1) <= Y1 <= U.*Del1, Pito2' - U.*(1-Del2) <= Y2 <= U.*Del2];
    
[…]

### 2013-02-18 project/report/paper.tex
almost ready

@@
 With this in mind, we present an aircraft EPS controller, or \emph{load management system (LMS)}, that guarantees safety while aiming to optimize quality of service metrics such as minimizing load shedding.
-Our controller uses a hierarchical control to combine an optimization-based model predictive LMS with a priority table based low-level LMS.
+Our controller is a hierarchical load management system that combines an optimization-based model predictive LMS with a priority table based low-level LMS.
 The fundamental contribution is in coupling the two controllers to work in a hierarchical fashion, where the optimization-based LMS looks forward and the low-level LMS guarantees safety in the short term.
@@
 
+~\\
 \section{Related Work}
@@
 \label{sec:optimizing-LMS}
-LMS systems that are based purely on priority tables are very simplistic, allowing them to frequently sample the system. 
+LMS systems that are based purely on priority tables are intended to be very simplistic, allowing them to frequently sample the system. 
 These LMS systems are thus more responsive to failures in the system.
-However, pure priority table solutions tend to shed more loads than necessary, to use more generators than necessary, 
+However, pure priority table solutions tend to shed more loads than necessary and to use more generators than necessary.
+It is possible to design a more intelligent low level system. However, more intelligent systems will tend to be more
+complex, making it more difficult to operate the system at the high frequencies desired in a low level system.
 %[and be ill-defined in their tradeoffs between generator allocation and load shedding].
@@
 \caption{\textbf{Load Priority Table Example, NS = nonsheddable, S = sheddable} 
-Loads lower in the list have lower priority while loads higher up in the list have
-higher priority. LL-LMS will shed load starting at the lower end of the list.}
+Loads higher in the list have higher shedding priority while loads lower in the 
+list have lower priority. LL-LMS will shed load starting at the top of the list.}
 \label{T:shed}
@@
 Because failures are rare, the LL-LMS will spend most of the time actuating the system based on the Optimizing LMS's advice. 
-This Hierarchical LMS is optimal because the Optimizing LMS is optimal and is safe because the LL-LMS is safe.
+Therefore in normal operating conditions, the Hierarchical LMS will be optimal because the system is 
+essentially being controlled by the Optimizing LMS. The only time the system is using a suboptimal
+configuration is when there is a failure and this is only temporary because eventually the Optimizing
+LMS will have readjusted for the failures.
 
@@
 Instead of having the Optimizing LMS system read the sensors directly as it would in a system that relies solely on an Optimizing LMS, it will instead receive its system status information from the LL-LMS.
-%This communication strategy saves on communication links between sensors and the Optimizing LMS and reuses the communication link between the LL-LMS and the Optimizing LMS. 
-This communication strategy allows the LL-LMS to have the final word on selecting a system configuration.
+This communication strategy saves on communication links between sensors and the Optimizing LMS and reuses the communication pathway between the LL-LMS and the Optimizing LMS. 
+%This communication strategy allows the LL-LMS to have the final word on selecting a system configuration.
 %We also use the Optimizing LMS to perform live optimizations of the system. 
@@
 Notice that there is an arrow going from the the LL-LMS to the Optimizing LMS because the LL-LMS is responsible for providing the Optimizing LMS with information of the status of the generators.
+Alternatively, we could have an additional arrow going from the sensors into the Optimizing LMS, but this is wasteful.
+There is no need for the Optimizing LMS to directly sample the sensors.
+We can instead
+use the existing communication path between the Optimzing LMS and the LL-LMS to send sensor data to the Optimzing LMS.
 The generator usage and load shedding that the LL-LMS will actuate on will be either the advice provided by the Optimizing LMS
@@
 Sheddable loads (W) & Shed priority \\ \hline
-60,000 & 1 (least preferred to shed) \\
+40,000 & 1 (most preferred to shed) \\ \hline
 60,000 & 2 (middle priority to shed) \\ 
-40,000 & 3 (most preferred to shed) \\ \hline
+60,000 & 3 (least preferred to shed) \\
 \end{tabular}
@@
 When presented with this workload, the LL-LMS purely applies the priority table, without any optimization or branch-and-bound to investigate ideal combinations of tasks to shed. 
-Thus, while simply shedding Load 2 would be sufficient to get system under its power cap, the LL-LMS walks up the priority table and sheds Loads 2 and 3.
+Thus, while simply shedding Load 2 would be sufficient to get system under its power cap, the LL-LMS walks down the priority table and sheds Loads 1 and 2 because it must
+obey the priority table. On the other hand, if loads 1 and 2 had equal priority, it would be possible for the LL-LMS to only shed Load 2.
 
@@
 Sheddable loads (W) & Shed priority \\ \hline
-40,000 & 1 (least preferred to shed) \\
+{\bf 20,000} & 1 (most preferred to shed) \\ \hline
 40,000 & 2 (middle priority to shed) \\ 
-{\bf 20,000} & 3 (most preferred to shed) \\ \hline
+40,000 & 3 (least preferred to shed) \\
 \end{tabular}
@@
 Sheddable loads (W) & Shed priority \\ \hline
-40,000 & 1 (least preferred to shed) \\
+{\bf 40,000} & 1 (most preferred to shed) \\ \hline
 40,000 & 2 (middle priority to shed) \\ 
-{\bf 40,000} & 3 (most preferred to shed) \\ \hline
+40,000 & 3 (least preferred to shed) \\
 \end{tabular}
@@
 Sheddable loads (W) & Shed priority \\ \hline
-40,000 & 1 (least preferred to shed) \\
+20,000 & 1 (most preferred to shed) \\ \hline
 40,000 & 2 (middle priority to shed) \\ 
-20,000 & 3 (most preferred to shed) \\ \hline
+40,000 & 3 (least preferred to shed) \\
 \end{tabular}

### 2013-02-18 project/refactored_mehdi_code/OptProb_Linear_Function.m
removed battery cap -- works well. (mehdi's idea). starting at a batteryMin that I've defined, and staying above that min

@@ -51,6 +51,7 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     Overflow1=sdpvar(1,Nt,'full'); Overflow2 = sdpvar(1,Nt,'full');
     wastePower1=sdpvar(1,Nt,'full'); wastePower2=sdpvar(1,Nt,'full'); %Power not used by loads
     BETA1 = sdpvar(1,Nt,'full'); BETA2 = sdpvar(1,Nt,'full'); %cumulative battery charge. (lowercase Beta is per-timestep change in charge)
+    %TODO: convert Watts * seconds to Wh for printouts
 
     % Constraints
     cons=[];
@@ -60,7 +61,10 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     %doing a running total of battery charge
     timestep = 1; % 1sec for now. using this to convert W to Wh for battery capacity
     batteryCapacity = 70000; %Wh
-    cons=[cons, 0 <= BETA1 <= batteryCapacity, 0 <= BETA2 <= batteryCapacity];
+    %cons=[cons, 0 <= BETA1 <= batteryCapacity, 0 <= BETA2 <= batteryCapacity];
+    %cons=[cons, 0 <= BETA1, 0 <= BETA2]
+    minBatteryLevel = 50000;
+    cons=[cons, minBatteryLevel <= BETA1, minBatteryLevel <= BETA2]; %min value 100, so that we can see it on graph
     cons=[cons, wastePower1 >= 0, wastePower2 >= 0]; %loads can't siphon imaginary power by making wastePower negative
 
     for i=1:Nl-1
@@ -72,17 +76,18 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     x=1:1:100;
     xi=0:N/(Nt-1):N; xi(1)=1;  % 0:10:100
 
-    cons=[cons, sum(C1.*interp1(x,Ls1',xi)',1) + sum(interp1(x,Lns1',xi),2)' == sum(Y1,1) - wastePower1];   %\sum cji(t)*lji(t)= \sum \deta_ji *P_source_i - Betaj
[…]

### 2013-02-19 project/refactored_mehdi_code/OptProb_Linear_Function.m
batt

@@ -48,8 +48,8 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     Pito1=sdpvar(Nt,Ns,'full');         Pito2=sdpvar(Nt,Ns,'full'); %Pito1(t,g) = "amount of pwr delivered by generator g to bus 1 at time t"
 
     %new decision variables for battery overflow
-    Overflow1=sdpvar(1,Nt,'full'); Overflow2 = sdpvar(1,Nt,'full');
-    wastePower1=sdpvar(1,Nt,'full'); wastePower2=sdpvar(1,Nt,'full'); %Power not used by loads
+    %Overflow1=sdpvar(1,Nt,'full'); Overflow2 = sdpvar(1,Nt,'full');
+    %wastePower1=sdpvar(1,Nt,'full'); wastePower2=sdpvar(1,Nt,'full'); %Power not used by loads
     BETA1 = sdpvar(1,Nt,'full'); BETA2 = sdpvar(1,Nt,'full'); %cumulative battery charge. (lowercase Beta is per-timestep change in charge)
     %TODO: convert Watts * seconds to Wh for printouts
 
@@ -63,9 +63,11 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     batteryCapacity = 70000; %Wh
     %cons=[cons, 0 <= BETA1 <= batteryCapacity, 0 <= BETA2 <= batteryCapacity];
     %cons=[cons, 0 <= BETA1, 0 <= BETA2]
-    minBatteryLevel = 50000;
+    startBatteryLevel = 0;
+    minBatteryLevel = 20000; %afterthe tMinBatteryLevel-th timestep
+    tMinBatteryLevel = 10; %first timestep to take minBatteryLevel into account
     cons=[cons, minBatteryLevel <= BETA1, minBatteryLevel <= BETA2]; %min value 100, so that we can see it on graph
-    cons=[cons, wastePower1 >= 0, wastePower2 >= 0]; %loads can't siphon imaginary power by making wastePower negative
+    %cons=[cons, wastePower1 >= 0, wastePower2 >= 0]; %loads can't siphon imaginary power by making wastePower negative
 
     for i=1:Nl-1
[…]

### 2013-02-19 project/refactored_mehdi_code/OptProb_Linear_Function.m
code cleanup

@@ -48,8 +48,6 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     Pito1=sdpvar(Nt,Ns,'full');         Pito2=sdpvar(Nt,Ns,'full'); %Pito1(t,g) = "amount of pwr delivered by generator g to bus 1 at time t"
 
     %new decision variables for battery overflow
-    %Overflow1=sdpvar(1,Nt,'full'); Overflow2 = sdpvar(1,Nt,'full');
-    %wastePower1=sdpvar(1,Nt,'full'); wastePower2=sdpvar(1,Nt,'full'); %Power not used by loads
     BETA1 = sdpvar(1,Nt,'full'); BETA2 = sdpvar(1,Nt,'full'); %cumulative battery charge. (lowercase Beta is per-timestep change in charge)
     %TODO: convert Watts * seconds to Wh for printouts
 
@@ -62,12 +60,10 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     timestep = 1; % 1sec for now. using this to convert W to Wh for battery capacity
     batteryCapacity = 70000; %Wh
     %cons=[cons, 0 <= BETA1 <= batteryCapacity, 0 <= BETA2 <= batteryCapacity];
-    %cons=[cons, 0 <= BETA1, 0 <= BETA2]
     startBatteryLevel = 0;
     minBatteryLevel = 20000; %afterthe tMinBatteryLevel-th timestep
     tMinBatteryLevel = 10; %first timestep to take minBatteryLevel into account
     cons=[cons, minBatteryLevel <= BETA1, minBatteryLevel <= BETA2]; %min value 100, so that we can see it on graph
-    %cons=[cons, wastePower1 >= 0, wastePower2 >= 0]; %loads can't siphon imaginary power by making wastePower negative
 
     for i=1:Nl-1
         cons=[cons, C1(i,:) <= C1(i+1,:), C2(i,:) <= C2(i+1,:)];
@@ -97,15 +93,11 @@ function [C1 C2 Del1 Del2 Beta1 Beta2 Y1 Y2 alpha Pito1 Pito2 ] = OptProb_Linear
     obj = obj + sum(Gamma1 * (1-C1)) + sum (Gamma2 * (1-C2));
     obj = obj + sum(Lambda1 * Del1) + sum(Lambda2 * Del2);
[…]
